---
title: "Regression and Other Stories: Parabola"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Demonstration of using Stan for optimization. See Appendix A in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstan)

# Parameters
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# A Computing in R

## A.7 Some R programming

### Opimization

#### Finding the peak of a parabola

Plot of $y = 15 + 10 x - 2 x^2$.

```{r, fig.asp=0.75}
v <- 
  tibble(
    x = seq_range(c(-2, 5)),
    y = 15 + 10 * x - 2 * x^2
  )

v %>% 
  ggplot(aes(x, y)) +
  geom_line() +
  geom_point(data = tibble(x = 2.5, y = 27.5)) +
  labs(
    title = expression(paste("Plot of ", y == 15 + 10 * x - 2 * x^2)),
    subtitle = "Maximum is at x = 2.5, y = 27.5"
  )
```

Stan model for function $15 + 10 x - 2 x^2$.

```{r}
model_code =
"
parameters {
  real x;
}
model {
  target += 15 + 10 * x - 2 * x^2;
}
"
```

Compile the Stan function and optimize it.

```{r, message=FALSE, warning=FALSE, error=FALSE, results=FALSE}
set.seed(245)

fit <- 
  stan_model(model_code = model_code) %>% 
  optimizing()
```

```{r}
fit
```

The output labeled `par` is the value of `x` at which the function is optimized, and the output labeled `value` is the target function at the optimum. A `return_code` of 0 corresponds to the optimizer running with no problem, and you can ignore `theta_tilde`.

