---
title: "Regression and Other Stories: Pew"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Miscellaneous analyses using raw Pew data. See Chapter 2 in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(haven)

# Parameters
  # Pew data on political ideology and party identification
file_pew <- here::here("Pew/data/pew_research_center_june_elect_wknd_data.dta")
  # Common code
file_common <- here::here("_common.R")
  
#===============================================================================

# Run common code
source(file_common)
```

# 2 Data and measurement

## 2.1 Examining where data come from

### Details of measurement can be important

Data

```{r}
pew <- 
  read_dta(file_pew) %>% 
  mutate(across(where(is.labelled), as_factor))

glimpse(pew)
```

Income.

```{r}
pew %>% 
  count(income, income2)
```

We'll use `income`, since it provides greater resolution at high incomes. We'll format `income` for plotting and recode "dk/refused" as `NA`.

```{r}
income_recode <- 
  levels(pew$income) %>% 
  set_names(
    levels(pew$income) %>% 
      str_replace("less", "Less") %>% 
      str_replace("-", " - ")
  )

pew <- 
  pew %>% 
  mutate(income = na_if(income, "dk/refused") %>% fct_recode(!!! income_recode))
```

Party identification.

```{r}
pew %>% 
  count(party, partyln, sort = TRUE)
```

We'll combine `party` and `partyln` into one variable `party_5`

```{r}
pew <- 
  pew %>% 
  mutate(
    party_5 =
      case_when(
        party == "democrat" ~ "Democrat",
        party == "republican" ~ "Republican",
        partyln == "lean democrat" ~ "Lean Democrat",
        partyln == "lean republican" ~ "Lean Republican",
        party == "independent" ~ "Independent",
        TRUE ~ NA_character_
      ) %>% 
      fct_relevel(
        "Democrat",
        "Lean Democrat",
        "Independent",
        "Lean Republican",
        "Republican"
      )
  )
```

```{r}
pew %>% 
  count(party, partyln, party_5, sort = TRUE)
```

Political ideology.

```{r}
pew %>% 
  count(ideo)
```

Again, we'll format `ideo` for plotting and recode "dk/refused" as `NA`.

```{r}
ideo_recode <- 
  levels(pew$ideo) %>% 
  set_names(levels(pew$ideo) %>% str_to_title())

pew <-  
  pew %>% 
  mutate(ideo = na_if(ideo, "dk/refused") %>% fct_recode(!!! ideo_recode))
```

Self-declared political ideology by income in 2008.

```{r, fig.asp=0.8}
v <- 
  pew %>% 
  drop_na(income, ideo) %>% 
  group_by(income, ideo) %>% 
  summarize(n = sum(weight)) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

v %>% 
  ggplot(aes(income, prop, group = ideo)) +
  geom_line(aes(color = ideo)) +
  ggrepel::geom_text_repel(
    aes(label = ideo),
    data = . %>% slice_max(order_by = income, n = 1),
    nudge_x = 0.2,
    direction = "y",
    hjust = 0
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.35))) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  coord_cartesian(ylim = c(0, NA)) +
  theme(
    axis.text.x = element_text(angle = -45, hjust = 0),
    legend.position = "none"
  ) +
  labs(
    title = "Self-declared political ideology by income in 2008",
    x = "Income",
    y = "Percentage of income group"
  )
```

Self-declared party identification by income in 2008.

```{r, fig.asp=0.8}
v <- 
  pew %>% 
  drop_na(income, party_5) %>% 
  group_by(income, party_5) %>% 
  summarize(n = sum(weight)) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

v %>% 
  ggplot(aes(income, prop, group = party_5)) +
  geom_line(aes(color = party_5)) +
  ggrepel::geom_text_repel(
    aes(label = party_5),
    data = . %>% slice_max(order_by = income, n = 1),
    nudge_x = 0.2,
    direction = "y",
    hjust = 0
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.35))) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_color_discrete(direction = -1) +
  coord_cartesian(ylim = c(0, NA)) +
  theme(
    axis.text.x = element_text(angle = -45, hjust = 0),
    legend.position = "none"
  ) +
  labs(
    title = "Self-declared party identification by income in 2008",
    x = "Income",
    y = "Percentage of income group"
  )
```

