---
title: "Regression and Other Stories: Electric Company"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Simple analysis of incentives data. See Chapter 19 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Incentives data
file_incentives <- here::here("Incentives/data/incentives.csv") 
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 19 Causal inference using direct regression

## 19.5 Challenges of interpreting regression coefficients as treatment effects

Data from a meta-analysis of studies of incentives in sample surveys.

```{r, message=FALSE}
incentives <- read_csv(file_incentives)

incentives
```

Fit linear regression.

```{r}
set.seed(447)

fit <- 
  stan_glm(
    rr_diff ~ value + prepay + gift + burden,
    data = incentives,
    refresh = 0
  )

print(fit, digits = 2)
```

The above coefficients should not be directly interpreted as causal effects. Although incentive conditions were assigned randomly _within_ each experiment, the differences in the conditions were not assigned at random _between_ experiments. Thus, when comparing incentives implemented in different surveys, what we have is an observational study.
