---
title: "Regression and Other Stories: Beauty and Teaching Quality"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

See Chapter 10 in Regression and Other Stories.

Hamermesh and Parker (2005) data on student evaluations of
instructors’ beauty and teaching quality for several courses at the
University of Texas. See Chapter 10 in Regression and Other
Stories.

Hamermesh, D. S., and Parker, A. M. (2005).  Beauty in the
classroom: Instructors' pulchritude and putative pedagogical
productivity.  Economics of Education Review, 24:369-376.
([Preprint](https://utexas.app.box.com/s/nl3ft5e1hunsnfomu87t))

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Course evaluation and instructor data
file_beauty <- here::here("Beauty/data/beauty.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 10 Linear regression and multiple predictions

## 10.11 Exercises

Data

```{r, message=FALSE}
beauty <- 
  file_beauty %>% 
  read_csv() %>% 
  mutate(across(c(female, minority, nonenglish, lower, course_id), factor))
```

#### 1D Exporatory data analysis

```{r}
glimpse(beauty)
```

```{r}
summary(beauty)
```

`eval`

```{r}
beauty %>% 
  ggplot(aes(eval)) +
  geom_bar()
```

Most evaluations are between 3 and 5 with a mean and median of 4.

`beauty`

```{r}
beauty %>% 
  ggplot(aes(beauty)) +
  geom_histogram(binwidth = 0.2)
```

`beauty` appears to have been normalized so that the mean is approximately 0.

`age`

```{r}
beauty %>% 
  ggplot(aes(age)) +
  geom_histogram(binwidth = 5, boundary = 0)
```

A fairly uniform distribution between ages 30 and 65, with mean and median around 48.

`female`

```{r}
beauty %>% 
  count(female) %>% 
  mutate(prop = n / sum(n))
```

About 42% of the instructors are female.

`minority`

```{r}
beauty %>% 
  count(minority) %>% 
  mutate(prop = n / sum(n))
```

About 14% of the instructors are from a minority.

`nonenglish`

```{r}
beauty %>% 
  count(nonenglish) %>% 
  mutate(prop = n / sum(n))
```

Only about 6% of the instructors are non-native English speakers.

`lower`

```{r}
beauty %>% 
  count(lower) %>% 
  mutate(prop = n / sum(n))
```

About 34% of the classes are lower division.

`couse_id`

```{r}
beauty %>% 
  count(course_id, sort = TRUE) %>% 
  mutate(prop = n / sum(n))
```

66% of courses have a `course_id` 0. It's hard to imagine 306 offerings of the same course. Perhaps 0 indicates `NA`. In any case, this does not appear to be a useful variable for modeling.

#### 2D Exporatory data analysis

##### Relationship of variables to `eval`

```{r}
beauty %>% 
  ggplot(aes(beauty, eval)) +
  geom_count() +
  geom_smooth(method = "loess", formula = "y ~ x")
```

There appears to be a slight positive relationship between `beauty` and `eval`.

```{r}
beauty %>% 
  ggplot(aes(age, eval)) +
  geom_count() +
  geom_smooth(method = "loess", formula = "y ~ x")
```

There does not appear to be much of a relationship between `age` and `eval`.

```{r}
beauty %>% 
  ggplot(aes(female, eval)) +
  geom_boxplot()
```

On average, female instructors receive lower evaluations.

```{r}
beauty %>% 
  ggplot(aes(minority, eval)) +
  geom_boxplot()
```

On average, minority instructors receive lower evaluations.

```{r}
beauty %>% 
  ggplot(aes(nonenglish, eval)) +
  geom_boxplot()
```

On average, instructors who are non-native English speakers receive lower evaluations.

```{r}
beauty %>% 
  ggplot(aes(lower, eval)) +
  geom_boxplot()
```

On average, lower division classes receive slightly higher evaluations.

##### Relationship of variables to `beauty`

```{r}
beauty %>% 
  ggplot(aes(age, beauty)) +
  geom_count() +
  geom_smooth(method = "loess", formula = "y ~ x")
```

There appears to be a negative relationship between `age` and `beauty`.

```{r}
beauty %>% 
  ggplot(aes(female, beauty)) +
  geom_boxplot()
```

Female instructors receive higher beauty scores.

```{r}
beauty %>% 
  ggplot(aes(minority, beauty)) +
  geom_boxplot()
```

Minority instructors receive roughly the same beauty scores as white instructors.

```{r}
beauty %>% 
  ggplot(aes(nonenglish, beauty)) +
  geom_boxplot()
```

Instructors who are non-native English speakers receive slightly higher beauty scores.

#### Linear regressions

Fit linear regression with predictor: `beauty`

```{r}
set.seed(616)

fit_1 <- stan_glm(eval ~ beauty, data = beauty, refresh = 0)

print(fit_1, digits = 2)
```

For instructors with the average beauty score of 0, the predicted teaching evaluation score is 4.0. The relationship of beauty score to predicted teaching evaluation score is slightly positive with a slope of 0.13.

Average teaching evaluation vs. beauty score.

```{r}
v <- 
  tibble(beauty = seq_range(beauty$beauty)) %>% 
  predictive_intervals(fit = fit_1)

v %>% 
  ggplot(aes(beauty)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), alpha = 0.25) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), alpha = 0.5) +
  geom_line(aes(y = .pred)) +
  geom_count(aes(y = eval), data = beauty) +
  theme(legend.position = "bottom") +
  labs(
    title = "Average teaching evaluation vs. beauty score",
    subtitle = "With 50% and 90% predictive intervals",
    x = "Beauty score",
    y = "Average teaching evaluation",
    size = "Count"
  )
```

There is considerable uncertainty in the relationship between the beauty score and the average teaching evaluation.

Fit linear regression with predictors: `beauty`, `female`

```{r}
set.seed(616)

fit_2 <- 
  stan_glm(eval ~ beauty + female + beauty:female, data = beauty, refresh = 0)

print(fit_2, digits = 2)
```

For instructors with the average beauty score of 0, the predicted teaching evaluation score is 4.1 for men and 3.9 for women. The relationship of beauty score to predicted teaching evaluation score is more positive for men, with a slope of 0.20 for men and 0.09 for women.

Average teaching evaluation vs. beauty score and sex.

```{r}
lines <- 
  tribble(
    ~female, ~intercept, ~slope,
    0, coef(fit_2)[["(Intercept)"]], coef(fit_2)[["beauty"]],
    1, 
      coef(fit_2)[["(Intercept)"]] + coef(fit_2)[["female1"]],
      coef(fit_2)[["beauty"]] + coef(fit_2)[["beauty:female1"]]
  ) %>% 
  mutate(across(female, factor))

beauty %>% 
  ggplot(aes(beauty, eval, color = female)) +
  geom_count() +
  geom_abline(
    aes(slope = slope, intercept = intercept, color = female),
    data = lines
  ) +
  scale_color_discrete(breaks = 1:0, labels = c("Female", "Male"), direction = -1) +
  theme(legend.position = "bottom") +
  labs(
    title = "Average teaching evaluation vs. beauty score and sex",
    x = "Beauty score",
    y = "Average teaching evaluation",
    color = "Sex",
    size = "Count"
  )
```

Fit linear regression with predictors: `beauty`, `female`, `minority`, `nonenglish`, `lower`

```{r}
set.seed(616)

fit_3 <- 
  stan_glm(
    eval ~ beauty + female + beauty:female + minority + nonenglish + lower,
    data = beauty,
    refresh = 0
  )

print(fit_3, digits = 2)
```

A male, white, native English speaking instructor, not teaching a lower division course, and with the average beauty score of 0, has the baseline predicted teaching evaluation score of 4.09. Compared to this baseline, for other instructors with the average beauty score of 0, the predicted teaching evaluation score for:

* Women is 0.19 lower.
* Minority instructors is 0.04 lower.
* Non-native English speaking instructors is 0.29 lower.
* Instructors of lower division classes is 0.09 higher.

The relationship of beauty score to predicted teaching evaluation score is more positive for men, with a slope of 0.20 for men and 0.09 for women.

Regression residuals vs. predicted values.

```{r}
v <- 
  tibble(
    pred = predict(fit_3),
    resid = residuals(fit_3)
  )

v %>% 
  ggplot(aes(pred, resid)) +
  geom_hline(yintercept = 0, color = "white", size = 2) +
  geom_count() +
  theme(legend.position = "bottom") +
  labs(
    title = "Regression residuals vs. predicted values",
    x = "Predicted value",
    y = "Residual",
    size = "Count"
  )
```

The residuals appear to be evenly distributed about 0 with no patterns.

