---
title: "Regression and Other Stories: Earnings"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Bootstrapping to simulate the sampling distribution. See Chapter 5 in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Earnings data
file_earnings <- here::here("Earnings/data/earnings.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 5 Simulation

## 5.4 Bootstrapping to simulate a sampling distribution

Data

```{r, message=FALSE}
earnings <- 
  file_earnings %>% 
  read_csv() %>% 
  mutate(
    sex = 
      case_when(
        male == 0 ~ "Female",
        male == 1 ~ "Male",
        TRUE ~ NA_character_
      )
  ) %>% 
  select(age, sex, earn)
```

EDA

```{r}
earnings %>% 
  select(age, earn) %>% 
  summary()

earnings %>% 
  count(sex)

earnings %>%
  group_by(sex) %>% 
  summarize(across(earn, list(median = median, mean = mean)))
```

Median of women's earnings divided by the median of men's earnings.

```{r}
earnings %>% 
  summarize(
    earnings_ratio = median(earn[sex == "Female"]) / median(earn[sex == "Male"])
  ) %>% 
  pull(earnings_ratio)
```

Earnings ratio of bootstrap sample.

```{r}
earnings_ratio <- function() {
  earnings %>% 
    select(sex, earn) %>% 
    slice_sample(prop = 1, replace = TRUE) %>% 
    summarize(
      earnings_ratio = 
        median(earn[sex == "Female"]) / median(earn[sex == "Male"])
    ) %>% 
    pull(earnings_ratio)
}
```

A set of bootstrap simulations.

```{r}
set.seed(194)

n_sims <- 1e4

earnings_ratio_boot <- 
  tibble(earnings_ratio = map_dbl(seq_len(n_sims), ~ earnings_ratio()))
```

Distribution of bootstrap earnings ratios.

```{r}
earnings_ratio_boot %>% 
  ggplot(aes(earnings_ratio)) +
  geom_histogram(binwidth = 0.01) +
  labs(
    title = "Distribution of bootstrap earnings ratios",
    x = "Earnings ratio",
    y = "Count"
  )
```

```{r}
earnings_ratio_boot %>% 
  count(earnings_ratio, sort = TRUE) %>% 
  mutate(prop_cum = cumsum(n) / sum(n))
```

Because of the nature of the data, one value accounts for 70% of the earnings ratios in the bootstrap samples and the top three account for 92%.

Standard deviation of bootstrap earnings ratios.

```{r}
sd(earnings_ratio_boot$earnings_ratio)
```

