---
title: "Regression and Other Stories: Earnings data"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Read in and prepare earnings data. See Appendix A in Regression and Other
Stories.

Source: Ross, Catherine E. Work, Family, and Well-Being in the United States, 1990. 1996-06-10. [https://doi.org/10.3886/ICPSR06666.v1](https://doi.org/10.3886/ICPSR06666.v1)

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Earnings data
file_earnings <- here::here("Earnings/data/wfw90.dat")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# A Computing in R

## A.6 Working with messy data

### Reading in survey data, one question at a time

Read in survey data.

```{r}
earnings <- 
  file_earnings %>% 
  read_fwf(
    col_positions = 
      fwf_cols(
        height_feet = 144,
        height_inches = c(145, 146),
        weight = c(147, 149),
        earn_exact = c(203, 208),
        earn2 = c(209, 210),
        sex = 219
      ),
    col_types = cols(.default = col_double())
  )
```

### Cleaning data within R

1. Look at the data.
2. Identify errors or missing data.

```{r}
earnings
```

From the first rows, it appears as though for each observation only one of `earn_exact` or `earn2` is non-`NA`. Let's see if this holds for all observations.

```{r}
earnings %>% 
  count(!is.na(earn_exact), !is.na(earn2))
```

It does.

```{r}
summary(earnings)
```

`height_feet` has values of 9 feet, `height_inches` has values of 99 inches, and `weight` has values of 999 pounds. From the data codebook, we see that the missing data codes are:

* Don't know: 8, 98, 998
* No response: 9, 99, 999

We'll recode these values as `NA`s.

```{r}
earnings <- 
  earnings %>% 
  mutate(
    height_feet = height_feet %>% na_if(8) %>% na_if(9),
    height_inches = height_inches %>% na_if(98) %>% na_if(99),
    weight = weight %>% na_if(998) %>% na_if(999)
  )

earnings %>% 
  select(height_feet, height_inches, weight) %>% 
  summary()
```

Let's look at the observations with `height_feet` of 7 feet.

```{r}
earnings %>% 
  filter(height_feet == 7)
```

The observation of someone 7 feet 7 inches tall who weighs 110 pounds is probably a data error, so we will recode the `height_feet` value of 7 as missing.

```{r}
earnings <- 
  earnings %>% 
  mutate(height_feet = height_feet %>% na_if(7))
```

<span>3</span>. Transform or combine raw data into summaries of interest.

Using the codebook, we'll first recode `sex`.

```{r}
earnings <- 
  earnings %>% 
   mutate(
    sex = 
      case_when(
        sex == 1 ~ "Male",
        sex == 2 ~ "Female",
        TRUE ~ NA_character_
      )
  )
```

We'll next create a combined variable `height`.

```{r}
earnings <- 
  earnings %>% 
  mutate(height = 12 * height_feet + height_inches)
```

Using the codebook, we'll create a new variable `earnings_approx` from `earn2`.

```{r}
earnings <- 
  earnings %>% 
  mutate(
    earn_approx =
      case_when(
        earn2 >= 90 ~ NA_real_,
        earn2 == 1 ~ 
          median(earn_exact[earn_exact > 100000], na.rm = TRUE) / 1000,
        TRUE ~ earn2
      )
  )
```

We'll finally create a combined earnings variable `earn`.

```{r}
earnings <- 
  earnings %>% 
  mutate(earn = if_else(!is.na(earn_exact), earn_exact, 1000 * earn_approx))
```

The new `earn` variable still has `r sum(is.na(earnings$earn))` missing values (out of `r nrow(earnings)` respondents in total) and is imperfect in various ways, but we have to make some choices when working with real data.

### Looking at the data

Sex distribution.

```{r}
earnings %>% 
  ggplot(aes(sex)) +
  geom_bar() +
  labs(title = "Sex distribution")
```

We have a greater number of women than men in the data, and a greater proportion than in the general adult population.

Height distributions.

```{r, fig.asp=0.75}
earnings %>% 
  drop_na(height) %>% 
  ggplot(aes(height)) +
  geom_bar() +
  facet_grid(rows = vars(sex)) +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  labs(title = "Height distributions")
```

There appears to be an excess of women who at 5 feet (60 inches) and 5 feet 6 inches (66 inches), and an excess of men who are 6 feet (72 inches), probably due to respondents rounding.

Weight distributions.

```{r, fig.asp=0.75}
earnings %>% 
  drop_na(weight) %>% 
  ggplot(aes(weight)) +
  geom_histogram(binwidth = 10, boundary = 0) +
  facet_grid(rows = vars(sex)) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(title = "Weight distributions")
```

Nothing remarkable.

Earnings distributions.

```{r, fig.asp=0.75}
earnings %>% 
  drop_na(earn) %>% 
  ggplot(aes(earn)) +
  geom_histogram(binwidth = 5e3, boundary = 0) +
  coord_cartesian(xlim = c(NA, 70e3)) +
  facet_grid(rows = vars(sex)) +
  scale_x_continuous(breaks = scales::breaks_width(10e3)) +
  labs(title = "Earnings distributions")
```

There are spikes in the lowest bin. Let's check to see if these are due to respondents with no earnings.

```{r}
v <- 
  earnings %>% 
  filter(earn == 0) %>% 
  count(sex)

v
```

`r sum(v$n)` respondents had no earnings, with `r v$n[v$sex == "Female"]` of these women.

Earnings distributions.

```{r}
earnings %>% 
  drop_na(earn) %>% 
  ggplot(aes(sex, earn)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(NA, 70e3)) +
  labs(title = "Earnings distributions")
```

Men have a higher median earnings.

```{r}
v <- 
  earnings %>% 
  group_by(sex) %>% 
  summarize(earn_median = median(earn, na.rm = TRUE))

v
```

The median earnings of men is `r format(v$earn_median[v$sex == "Male"] - v$earn_median[v$sex == "Female"])` higher.

