---
title: "Regression and Other Stories: Earnings"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Predict respondents' yearly earnings using survey data from
1990. See Chapter 15 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Seed
SEED <- 7783
  # Earnings data
file_earnings <- here::here("Earnings/data/earnings.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 15 Other generalized linear models

## 15.8 Going beyond generalized linear models

### Mixed discrete/continuous data

Data

```{r, message=FALSE}
earnings <- 
  file_earnings %>% 
  read_csv() %>% 
  mutate(
    sex = 
      case_when(
        male == 0 ~ "Female",
        male == 1 ~ "Male",
        TRUE ~ NA_character_
      ),
    earn_pos = earn > 0
  )

earnings %>% 
  select(height, sex, earn, earn_pos)
```

It can be appropriate to model a variable such as earnings in two steps: first a logistic regression for Pr(y > 0) fit to all the data, then a linear regression on log(y), fit just to the subset of the data for which y > 0.

We first fit a logistic regression to predict whether earnings are positive.

```{r}
fit_pos <- 
  stan_glm(
    earn_pos ~ height + sex,
    family = binomial(link = "logit"),
    data = earnings,
    refresh = 0,
    seed = SEED
  )

print(fit_pos, digits = 2)
```

We then fit a log regression model to the respondents with positive earnings.

```{r}
fit_log_2 <- 
  stan_glm(
    log(earn) ~ height + sex,
    data = earnings %>% filter(earn_pos),
    refresh = 0,
    seed = SEED
  )

print(fit_log_2, digits = 2)
```

Thus, for example, a 66-inch-tall woman has an estimated probability of

```{r}
v <- plogis(coef(fit_pos)[["(Intercept)"]] + coef(fit_pos)[["height"]] * 66)

v
```

or an `r round(100 * v)`% chance, of having positive income.

If her earnings are positive, their predicted value is

```{r}
exp(coef(fit_log_2)[["(Intercept)"]] + coef(fit_log_2)[["height"]] * 66)
```

Combining these gives a mixture of a spike at 0 and a lognormal distribution, which is most easily manipulated using simulations.

```{r}
set.seed(SEED)

new <- tibble(height = 66, sex = "Female")

pred_pos <- 
  posterior_predict(fit_pos, newdata = new) %>% 
  as.logical()
pred_log_2 <- 
  posterior_predict(fit_log_2, newdata = new) %>% 
  as.numeric()
pred <- if_else(pred_pos, exp(pred_log_2), 0)
```

Predicted earnings for 66-inch-tall women.

```{r, fig.asp=0.75}
tibble(pred) %>% 
  ggplot(aes(pred)) + 
  geom_histogram(binwidth = 2000, boundary = 0) +
  coord_cartesian(xlim = c(NA, 1e5)) +
  scale_x_continuous(labels = scales::label_comma()) +
  labs(
    title = "Predicted earnings for 66-inch-tall women",
    subtitle = "Excluding outliers",
    x = "Earnings",
    y = "Count"
  )
```

