---
title: "Regression and Other Stories: Height and weight"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Predict weight from height. See Chapters 9 and 10 in Regression
and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Earnings data
file_earnings <- here::here("Earnings/data/earnings.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 9 Prediction and Bayesian inference

## 9.2 Prediction and uncertainty: `predict()`, `posterior_linpred()`, and `posterior_predict()`

Data

```{r, message=FALSE}
earnings <- 
  file_earnings %>% 
  read_csv() %>% 
  mutate(
    sex = 
      case_when(
        male == 0 ~ "Female",
        male == 1 ~ "Male",
        TRUE ~ NA_character_
      )
  )

earnings %>% 
  select(height, weight, sex, ethnicity)
```

```{r}
mean(earnings$height)
```

### Simulating uncertainty for the linear predictor and new observations

Fit linear regression of weight against height.

The option `refresh = 0` suppresses the default Stan sampling progress output. This is useful for small data with fast computation. For more complex models and bigger data, it can be useful to see the progress.

```{r}
set.seed(733)

fit_1 <- stan_glm(weight ~ height, data = earnings, refresh = 0)

fit_1
```

Center heights.

```{r}
earnings <- 
  earnings %>% 
  mutate(height_c = height - 66)
```

Fit using centered heights.

```{r}
set.seed(733)

fit_2 <- stan_glm(weight ~ height_c, data = earnings, refresh = 0)

fit_2
```

New data for a 70-inch-tall person.

```{r}
new <- tibble(height_c = 4)
```

Point prediction.

```{r}
point_pred <- predict(fit_2, newdata = new)

as.double(point_pred)
```

Posterior simulations where variation is from posterior uncertainty in the coefficients.

```{r}
linpred <- posterior_linpred(fit_2, newdata = new)
```

Posterior simulations where variation is from posterior uncertainty in the coefficients and predictive uncertainty.

```{r}
set.seed(733)

pred <- posterior_predict(fit_2, newdata = new)
```

```{r}
preds <- tibble(linpred = as.double(linpred), pred = as.double(pred))

preds %>% 
  summarize(across(everything(), list(median = median, mad = mad)))
```

The medians of the two sets of posterior simulations are similar, but the median absolute deviations are very different.

Distributions of posterior predictions.

```{r}
preds %>% 
  ggplot(aes(y = stat(density))) +
  geom_freqpoly(aes(pred, color = "pred"), binwidth = 10) +
  geom_freqpoly(aes(linpred, color = "linpred"), binwidth = 0.2) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  scale_y_continuous(breaks = 0) +
  scale_color_discrete(
    breaks = c("linpred", "pred"),
    labels = c("Linear predictor", "Predicted value")
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Distributions of posterior predictions",
    x = "Weight (pounds)",
    y = NULL,
    color = NULL
  )
```

# 10 Linear regression with multiple predictions

## 10.4 Indicator variables

Data

```{r}
earnings %>% 
  select(height, weight, sex, ethnicity)
```

Fit linear regression of weight against height.

```{r}
set.seed(733)

fit_1 <- stan_glm(weight ~ height, data = earnings, refresh = 0)

fit_1
```

Predict the weight for a 66-inch-tall person.

```{r}
set.seed(733)

new <- tibble(height = 66)

pred <- posterior_predict(fit_1, newdata = new)
```

```{r, comment=""}
cat(
  str_glue(
    "The predicted weight for 66-inch-tall person is ",
    "{format(mean(pred), digits = 1, nsmall = 1)} pounds with a sd of ",
    "{format(sd(pred), digits = 1, nsmall = 1)}."
  )
)
```

### Centering a predictor

Center heights.

```{r}
earnings <- 
  earnings %>% 
  mutate(height_c = height - 66)
```

Fit using centered heights.

```{r}
set.seed(733)

fit_2 <- stan_glm(weight ~ height_c, data = earnings, refresh = 0)

fit_2
```

### Including a binary variable in a regression

Fit linear regression of weight against height and sex.

```{r}
set.seed(733)

fit_3 <- stan_glm(weight ~ height_c + sex, data = earnings, refresh = 0)

fit_3
```

Predict the weight for a 70-inch-tall woman.

```{r}
set.seed(733)

new <- tibble(height_c = 4, sex = "Female")

pred <- posterior_predict(fit_3, newdata = new)
```

```{r, comment=""}
cat(
  str_glue(
    "The predicted weight for 70-inch-tall woman is ",
    "{format(mean(pred), digits = 1, nsmall = 1)} pounds with a sd of ",
    "{format(sd(pred), digits = 1, nsmall = 1)}."
  )
)
```

Predict the weight for a 70-inch-tall man.

```{r}
set.seed(733)

new <- tibble(height_c = 4, sex = "Male")

pred <- posterior_predict(fit_3, newdata = new)
```

```{r, comment=""}
cat(
  str_glue(
    "The predicted weight for 70-inch-tall man is ",
    "{format(mean(pred), digits = 1, nsmall = 1)} pounds with a sd of ",
    "{format(sd(pred), digits = 1, nsmall = 1)}."
  )
)
```

### Using indicator variables for multiple levels of a categorical variable

```{r}
earnings %>% 
  count(ethnicity)
```

Fit linear regression of weight against height, sex, and ethnicity.

```{r}
set.seed(733)

fit_4 <- 
  stan_glm(weight ~ height_c + sex + ethnicity, data = earnings, refresh = 0)

fit_4
```

### Changing the baseline factor level

Make `"White"` the baseline factor level for `ethnicity`.

```{r}
earnings <- 
  earnings %>% 
  mutate(
    ethnicity = 
      factor(ethnicity, levels = c("White", "Black", "Hispanic", "Other"))
  )
```

Fit linear regression of weight against height, sex, and ethnicity with new baseline factor level.

```{r}
set.seed(733)

fit_5 <- 
  stan_glm(weight ~ height_c + sex + ethnicity, data = earnings, refresh = 0)

fit_5
```

Equivalently, create indicator variables for the four ethnic groups.

```{r}
earnings <- 
  earnings %>% 
  mutate(
    ethnicity_white = if_else(ethnicity == "White", 1, 0),
    ethnicity_black = if_else(ethnicity == "Black", 1, 0),
    ethnicity_hispanic = if_else(ethnicity == "Hispanic", 1, 0),
    ethnicity_other = if_else(ethnicity == "Other", 1, 0)
  )
```

Fit linear regression of weight against height, sex, and ethnicity expressed as indicator variables.

```{r}
set.seed(733)

fit_6 <- 
  stan_glm(
    weight ~ 
      height_c + sex + ethnicity_black + ethnicity_hispanic + ethnicity_other,
    data = earnings,
    refresh = 0
  )

fit_6
```

