---
title: "Regression and Other Stories: AgePeriodCohort"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r format(Sys.Date())`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Age-Period-Cohort - Demonstration of age adjustment to estimate
trends in mortality rates. See Chapter 2 in Regression and Other
Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Mortality for white non-Hispanics ages 45 - 54 from 1999 - 2018
file_mortality <- 
  here::here("AgePeriodCohort/data/white_mortality_45-54_1999-2018.txt")
  # Common code
file_common <- here::here("_common.R")
  
#===============================================================================

# Run common code
source(file_common)
```

# 2 Data and measurement

## 2.4 Data and adjustment: trends in mortality rates

Data

The dataset was downloaded from the CDC WONDER [Underlying Cause of Death](https://wonder.cdc.gov/ucd-icd10.html) database. It contains mortality data for white non-Hispanics ages 45 - 54, and breaks down mortality by year, age, gender, and census region. We will restrict attention to the years 1999 - 2013. The bottom of the dataset contains notes on the query and settings used to generate it. 

```{r, message=FALSE, warning=FALSE}
mortality <- 
  file_mortality %>% 
  read_tsv(
    col_types = 
      cols(
        Notes = col_character(),
        Year = col_double(),
        `Year Code` = col_double(),
        `Single-Year Ages` = col_character(),
        `Single-Year Ages Code` = col_double(),
        Gender = col_character(),
        `Gender Code` = col_character(),
        `Census Region` = col_character(),
        `Census Region Code` = col_character(),
        Deaths = col_double(),
        Population = col_double(),
        `Crude Rate` = col_double()
      )
  ) %>% 
  filter(is.na(Notes), Year >= 1999, Year <= 2013) %>%
  transmute(
    year = Year,
    age = `Single-Year Ages Code`,
    gender = Gender,
    region = str_extract(`Census Region`, "\\w+$"),
    deaths = Deaths,
    population = Population
  )

mortality
```

Unadjusted mortality for white non-Hispanics ages 45 - 54.

```{r}
mortality_unadjusted <- 
  mortality %>% 
  group_by(year) %>% 
  summarize(rate = sum(deaths) / sum(population) * 1e5)

mortality_unadjusted %>% 
  ggplot(aes(year, rate)) +
  geom_line() +
  labs(
    title = "Unadjusted mortality for white non-Hispanics ages 45 - 54",
    x = "Year",
    y = "Deaths per 100,000"
  )
```

Mean age for white non-Hispanics ages 45 - 54.

```{r}
v <- 
  mortality %>% 
  group_by(year) %>% 
  summarize(mean_age = weighted.mean(age, population))

v %>% 
  ggplot(aes(year, mean_age)) +
  geom_line() +
  labs(
    title = "Mean age for white non-Hispanics ages 45 - 54",
    x = "Year",
    y = "Mean age"
  )
```

A helper function to set the mortality rates for all years to that of a given year.

```{r}
set_rates <- function(.data, rate_year) {
  .data %>% 
    left_join(
      .data %>% 
        filter(year == rate_year) %>% 
        mutate(rate = deaths / population * 1e5) %>% 
        select(!c(year, deaths, population)),
      by = names(.data) %>% setdiff(c("year", "deaths", "population"))
    )
}
```

Mortality for white non-Hispanics ages 45 - 54.

```{r, fig.asp=0.7}
v <- 
  c(1999, 2013) %>% 
  set_names() %>% 
  map_dfr(~ set_rates(mortality, .), .id = "rate_year") %>% 
  group_by(rate_year, year) %>% 
  summarize(rate = weighted.mean(rate, population)) %>% 
  ungroup() %>% 
  bind_rows(
    mortality_unadjusted %>% 
      mutate(rate_year = "Unadjusted")
  )

v %>% 
  ggplot(aes(year, rate, color = rate_year)) +
  geom_line() +
  scale_color_manual(
    breaks = c("1999", "Unadjusted", "2013"),
    values = c("#F8766D", "black", "#00BFC4"),
    labels = c("1999 mortality rates", "Unadjusted", "2013 mortality rates")
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Mortality for white non-Hispanics ages 45 - 54",
    x = "Year",
    y = "Deaths per 100,000",
    color = NULL
  )
```

From 1999 to 2013, the unadjusted mortality rate increased from 382 to 415 deaths per 100,000, an increase of 33 deaths per 100,000. If the mortality rates for each age remained unchanged from the 1999 rates, the changing population distribution for the cohort would have increased its overall mortality rate from 382 to 398 deaths per 100,000, an increase of 17 deaths per 100,000. Thus the aging of the cohort played a substantial role in the increase of the unadjusted mortality.

A helper function to add the weights for age adjustment.

```{r}
add_weight <- function(.data, weight_year) {
  .data %>% 
    left_join(
      .data %>%
        filter(year == weight_year) %>% 
        select(!c(year, deaths), weight = population),
      by = names(.data) %>% setdiff(c("year", "deaths", "population"))
    )
}
```

Mortality for white non-Hispanics ages 45 - 54: Age-adjusted.

```{r, fig.asp=0.7}
v <- 
  c(1999, 2013) %>% 
  set_names() %>% 
  map_dfr(~ add_weight(mortality, .), .id = "weight_year") %>% 
  group_by(weight_year, year, age) %>% 
  summarize(
    across(c(deaths, population, weight), sum),
    rate = deaths / population * 1e5
  ) %>% 
  summarize(rate = weighted.mean(rate, weight)) %>% 
  ungroup() %>% 
  bind_rows(
    mortality_unadjusted %>% 
      mutate(weight_year = "Unadjusted")
  )

v %>% 
  ggplot(aes(year, rate, color = weight_year)) +
  geom_line() +
  scale_color_manual(
    breaks = c("1999", "Unadjusted", "2013"),
    values = c("#F8766D", "black", "#00BFC4"),
    labels = c("Age-adjusted 1999", "Unadjusted", "Age-adjusted 2013")
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Mortality for white non-Hispanics ages 45 - 54",
    x = "Year",
    y = "Deaths per 100,000",
    color = NULL
  )
```

As mentioned above, from 1999 to 2013 the unadjusted mortality rate increased from 382 to 415 deaths per 100,000, an increase of 33 deaths per 100,000. If the cohort had maintained the same age distribution in 2013 as in 1999, then the changing mortality rates each year would have increased the cohort's overall mortality rate from 382 to 400 deaths per 100,000, an increase of 18 deaths per 100,000.

To focus only on the changes in mortality and ignore the influence of the aging of the cohort, in the following we'll age adjust using the 1999 age distribution.

```{r}
mortality_1999 <- 
  mortality %>% 
  add_weight(weight_year = 1999)
```

Age-adjusted mortality for white non-Hispanics ages 45 - 54: By gender.

```{r, fig.asp=0.7}
v <- 
  mortality_1999 %>% 
  group_by(year, gender, age) %>% 
  summarize(
    across(c(deaths, population, weight), sum),
    rate = deaths / population * 1e5
  ) %>% 
  summarize(rate = weighted.mean(rate, weight)) %>% 
  ungroup()

v %>% 
  ggplot(aes(year, rate, color = gender)) +
  geom_line() +
  theme(legend.position = "bottom") +
  labs(
    title = "Age-adjusted mortality for white non-Hispanics ages 45 - 54",
    x = "Year",
    y = "Deaths per 100,000",
    color = NULL
  )
```

The mortality rate for men is substantially higher than that for women, but the genders have different trends. The rate for men in deaths per 100,000 was 484 in 1999, increased to a maximum of 515 in 2005, and then declined to 492 in 2013. The rate for women in deaths per 100,000 increased throughout the period, from 281 in 1999 to 309 in 2013. The combination of these two trends explains why the overall trend increased from 1999 to 2005 and then was stable afterwards. The decrease in the men's rate was roughly canceled by the increase in the women's rate.

Age-adjusted mortality for white non-Hispanics ages 45 - 54: By gender and region.

```{r, fig.asp=0.7}
v <- 
  mortality_1999 %>% 
  group_by(year, gender, region, age) %>% 
  summarize(
    across(c(deaths, population, weight), sum),
    rate = deaths / population * 1e5
  ) %>% 
  summarize(rate = weighted.mean(rate, weight)) %>% 
  ungroup()

v %>% 
  ggplot(aes(year, rate, color = region)) +
  geom_line() +
  facet_grid(cols = vars(gender)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Age-adjusted mortality for white non-Hispanics ages 45 - 54",
    x = "Year",
    y = "Deaths per 100,000",
    color = NULL
  )
```

We see a dramatic increase in mortality for women in the South, from 302 deaths per 100,000 in 1999 to 367 deaths per 100,000 in 2013.

