---
title: "Regression and Other Stories: Coop"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Coop - Example of hypothesis testing. See Chapter 4 in Regression
and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Votes in an election
file_votes <- here::here("Coop/data/Riverbay.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 4 Statistical inference

## 4.6 Example of hypothesis testing: 55,000 residents need your help!

Data

```{r, message=FALSE}
votes <- 
  file_votes %>% 
  read_csv(
    col_names = 
      c("name_1", "600", "1200", "2444", "3444", "4444", "5553", "name_2")
  ) %>% 
  pivot_longer(
    cols = !starts_with("name"),
    names_to = "voters",
    names_transform = list(voters = as.double),
    values_to = "votes"
  ) %>% 
  group_by(name_1) %>% 
  mutate(
    vote_prop_cum = votes / voters,
    vote_prop_new = 
      (votes - lag(votes, default = 0, order_by = voters)) / 
      (voters - lag(voters, default = 0, order_by = voters))
  ) %>% 
  ungroup()

votes
```

Candidates who received the most votes.

```{r}
voters <- 
  votes %>% 
  pull(voters) %>% 
  unique() %>% 
  sort()
voters_total <- max(voters)

votes %>% 
  filter(voters == voters_total) %>% 
  select(starts_with("name"), votes) %>% 
  arrange(desc(votes))
```

Top 8 candidates.

```{r}
candidates_top_8 <- 
  votes %>% 
  filter(voters == voters_total) %>% 
  slice_max(order_by = votes, n = 8) %>% 
  pull(name_2)
```

Cumulative vote percentages at different points in counting.

```{r}
votes_top_8 <- 
  votes %>% 
  filter(name_2 %in% candidates_top_8) %>% 
  mutate(name = ordered(name_2, levels = candidates_top_8))

votes_top_8 %>% 
  ggplot(aes(voters, vote_prop_cum)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  facet_wrap(facets = vars(name), ncol = 4) +
  theme(axis.text = element_text(size = rel(0.6))) +
  labs(
    title = "Cumulative vote percentages at different points in counting",
    subtitle = "For top 8 candidates",
    x = "Cumulative voters",
    y = "Cumulative vote percentage",
    caption = "Voters could vote for up to 6 candidates"
  )
```

Vote percentages of new votes at different points in counting.

```{r}
votes_top_8 %>% 
  ggplot(aes(voters, vote_prop_new)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  facet_wrap(facets = vars(name), ncol = 4) +
  theme(axis.text = element_text(size = rel(0.6))) +
  labs(
    title = "Vote percentages of new votes at different points in counting",
    subtitle = "For top 8 candidates",
    x = "Cumulative voters",
    y = "Vote percentage of new votes",
    caption = "Voters could vote for up to 6 candidates"
  )
```

Standard deviation of vote proportions of new votes for each candidate.

```{r}
v <- 
  votes %>% 
  group_by(name_1) %>% 
  summarize(
    votes_final = votes[voters == voters_total],
    vote_prop_new_sd = sd(vote_prop_new),
  )

voters_new <- voters - lag(voters, default = 0)
vote_prop_new_sd_expected <- 
  tibble(
    votes_final = seq_range(c(min(v$votes_final), max(v$votes_final))),
    vote_prop_final = votes_final / voters_total,
    vote_prop_new_sd = 
      map_dbl(vote_prop_final, ~ sqrt(mean(. * (1 - .) / voters_new)))
  )

ggplot(mapping = aes(votes_final, vote_prop_new_sd)) +
  geom_line(data = vote_prop_new_sd_expected) +
  geom_point(data = v) +
  scale_x_continuous(breaks = scales::breaks_width(500)) +
  labs(
    title = 
      "Standard deviation of vote proportions of new votes for each candidate",
    subtitle = 
      "Line is theoretical standard deviation expected under null hypothesis",
    x = "Total votes for candidate",
    y = "Standard deviation of vote proportions"
  )
```

