---
title: "Regression and Other Stories: ChileSchools"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Code and figures for ChileSchools example. See Chapter 21 in
Regression and Other Stories.

Data are from

- Chay, K. Y., McEwan, P. J., and Urquiola, M. (2005). The central
  role of noise in evaluating interventions that use test scores to
  rank schools. American Economic Review 95, 1237–1258.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Seed
SEED <- 733
  # Results of educational experiment
file_chile <- here::here("ChileSchools/data/chile.csv") 
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 21 Additional topics in causal inference

## 21.3 Regression discontinuity: known assignment mechanism but no overlap

### Example: The effect of an educational program on test scores in Chile

Data

```{r, message=FALSE}
chile_all <- read_csv(file_chile)

glimpse(chile_all)
```

Check consistency of `eligible` and `rule2`.

```{r}
chile_all %>% 
  count(eligible, rule2 < 0, rule2 == 0, rule2 > 0)
```

`eligible` accurately reflects whether `rule2` is greater than or less than 0.

#### Basic regression discontinuity analysis

Fit linear regression on all data.

```{r}
set.seed(SEED)

fit_1 <- stan_glm(read92 ~ eligible + rule2, data = chile_all, refresh = 0)

fit_1
```

Intent-to-treat effect of exposure to P-900 on schools in Chile: All data.

```{r, fig.asp=0.75}
coefs <- 
  tribble(
    ~eligible, ~intercept, ~slope,
    0, coef(fit_1)[["(Intercept)"]], coef(fit_1)[["rule2"]],
    1, coef(fit_1)[["(Intercept)"]] + coef(fit_1)[["eligible"]],
      coef(fit_1)[["rule2"]]
  )

plot <- function(data, coefs, all = FALSE, binned = FALSE) {
  segments <- 
    coefs %>% 
    mutate(
      x = if_else(eligible == 1, min(data$rule2), 0),
      xend = if_else(eligible == 1, 0, max(data$rule2)),
      y = intercept + slope * x,
      yend = intercept + slope * xend
    )

  size <- if_else(binned, 3, if_else(all, 0.75, 1.5))
  alpha <- if_else(binned, 1, if_else(all, 0.75, 1))
  breaks_width <- if_else(all, 10, 2)
  
  data %>%
    ggplot(aes(rule2, read92)) +
    geom_point(aes(color = factor(eligible)), size = size, alpha = alpha) +
    geom_vline(xintercept = 0) +
    geom_segment(aes(x, y, xend = xend, yend = yend), data = segments) +
    scale_x_continuous(breaks = scales::breaks_width(breaks_width)) +
    scale_color_discrete(
      breaks = c(1, 0),
      labels = c("Eligible", "Not eligible"),
      direction = -1
    ) +
    theme(legend.position = "bottom") +
    labs(
      title = "Intent-to-treat effect of exposure to P-900 on schools in Chile",
      x = "Assignment variable",
      color = "School eligibility"
    )
}

chile_all %>% 
  plot(coefs = coefs, all = TRUE) +
  labs(
    subtitle = "All data",
    y = "Post-test score"
  )
```

Data near cutoff.

```{r}
chile_5 <- 
  chile_all %>% 
  filter(abs(rule2) <= 5)
```

Fit linear regression on data near cutoff.

```{r}
set.seed(SEED)

fit_2 <- stan_glm(read92 ~ eligible + rule2, data = chile_5, refresh = 0)

fit_2
```

Intent-to-treat effect of exposure to P-900 on schools in Chile: Data near cutoff.

```{r, fig.asp=0.75}
coefs <- 
  tribble(
    ~eligible, ~intercept, ~slope,
    0, coef(fit_2)[["(Intercept)"]], coef(fit_2)[["rule2"]],
    1, coef(fit_2)[["(Intercept)"]] + coef(fit_2)[["eligible"]],
      coef(fit_2)[["rule2"]]
  )

chile_5 %>% 
  plot(coefs = coefs) +
  labs(
    subtitle = "Data near cutoff",
    y = "Post-test score"
  )
```

#### Including other pre-treatment variables as regression predictors

Fit linear regression on data near cutoff, including pre-treatment variables.

```{r}
set.seed(SEED)

fit_3 <- 
  stan_glm(
    read92 ~ eligible + rule2 + read88 + math88,
    data = chile_5,
    refresh = 0
  )

fit_3
```

Intent-to-treat effect of exposure to P-900 on schools in Chile: Data near cutoff.

```{r, fig.asp=0.75}
v <- 
  chile_5 %>% 
  mutate(
    read92 = 
      read92 - coef(fit_3)[["read88"]] * (read88 - mean(read88)) -
      coef(fit_3)[["math88"]] * (math88 - mean(math88))
  )

intercept_0 <- 
  coef(fit_3)[["(Intercept)"]] + 
  coef(fit_3)[["read88"]] * mean(chile_5$read88) +
  coef(fit_3)[["math88"]] * mean(chile_5$math88)

coefs <- 
  tribble(
    ~eligible, ~intercept, ~slope,
    0, intercept_0, coef(fit_3)[["rule2"]],
    1, intercept_0 + coef(fit_3)[["eligible"]], coef(fit_3)[["rule2"]]
  )

v %>% 
  plot(coefs = coefs) +
  labs(
    subtitle = "Data near cutoff",
    y = "Adjusted post-test score"
  )
```

Intent-to-treat effect of exposure to P-900 on schools in Chile: Data near cutoff binned

```{r, fig.asp=0.75}
n_bins <- 20

v <- 
  chile_5 %>% 
  mutate(
    read92 = 
      read92 - coef(fit_3)[["read88"]] * (read88 - mean(read88)) -
      coef(fit_3)[["math88"]] * (math88 - mean(math88)),
    bin = 
      cut(
        rule2,
        breaks = seq(-5, 5, length.out = n_bins + 1),
        include.lowest = TRUE
      )
  ) %>% 
  group_by(bin) %>% 
  summarize(
    eligible = as.double(min(rule2) < 0),
    rule2 = mean(rule2),
    read92 = mean(read92)
  )

v %>% 
  plot(coefs = coefs, binned = TRUE) +
  labs(
    subtitle = "Data near cutoff binned",
    y = "Adjusted post-test score"
  )
```

This plot is different from the one in the book. The adjustment above only uses the data near the cutoff that is used in the model. The plot in the book adjusts using all data.

#### Including interactions

Center and scale `read88` and `math88`.

```{r}
chile_5 <- 
  chile_5 %>% 
  mutate(
    read88_z = scale(read88) %>% as.double(),
    math88_z = scale(math88) %>% as.double()
  )
```

Fit linear regression with interaction.

```{r}
set.seed(SEED)

fit_4 <- 
  stan_glm(
    read92 ~ eligible + rule2 + read88_z + math88_z + eligible:read88_z,
    data = chile_5,
    refresh = 0
  )

fit_4
```

The results are again different from those in the book, since we have centered and scaled only using the data near the cutoff that is used in the model.

In this case, the available data are clearly too noisy to estimate this interaction.

#### Regression fit to all the data

Fit linear regression on all data, including pre-treatment variables.

```{r}
set.seed(SEED)

fit_5 <- 
  stan_glm(
    read92 ~ eligible + rule2 + read88 + math88,
    data = chile_all,
    refresh = 0
  )

fit_5
```

