---
title: "Regression and Other Stories: Pearson and Lee Heights"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

The heredity of height. Published in 1903 by Karl Pearson and Alice
Lee. See Chapter 6 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Heights of mothers and daughters from Pearson and Lee (1903)
file_heights <- here::here("PearsonLee/data/Heights.txt")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 6 Background on regression modeling

## 6.4 Historical origins of regression

### Fitting the model in R

Data

```{r}
heights <- 
  file_heights %>% 
  read.table(header = TRUE) %>% 
  as_tibble()

heights
```

Fit linear regression.

```{r}
set.seed(201)

fit <- stan_glm(daughter_height ~ mother_height, data = heights, refresh = 0)

print(fit, digits = 2)
```

Mothers' and adult daughters' heights.

```{r, fig.asp=1}
heights %>% 
  ggplot(aes(mother_height, daughter_height)) + 
  geom_count() +
  coord_fixed() +
  labs(
    title = "Mothers' and adult daughters' heights",
    x = "Mother's height (inches)",
    y = "Adult daughter's height (inches)"
  )
```

Mothers' and adult daughters' heights: With fitted regression line and average of data.

```{r, fig.asp=1}
heights_mean <- 
  heights %>% 
  summarize(across(everything(), mean))

intercept <- coef(fit)[["(Intercept)"]]
slope <- coef(fit)[["mother_height"]]

heights %>% 
  ggplot(aes(mother_height, daughter_height)) + 
  geom_count() +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point(data = heights_mean, color = "red", size = 2) +
  coord_fixed() +
  labs(
    title = "Mothers' and adult daughters' heights",
    subtitle = "With fitted regression line and average of data",
    x = "Mother's height (inches)",
    y = "Adult daughter's height (inches)"
  )
```

Fitted regression line and average of data.

```{r, fig.asp=1}
eqn_1 <- 
  str_glue(
    "y = {format(heights_mean$daughter_height, digits = 1, nsmall = 1)} + ",
    "{format(slope, digits = 3, nsmall = 3)} ",
    "(x - {format(heights_mean$mother_height, digits = 1, nsmall = 1)})"
  )
eqn_2 <- 
  str_glue(
    "y = {format(intercept, digits = 1, nsmall = 1)} + ",
    "{format(slope, digits = 3, nsmall = 3)} x"
  )
eqns <- str_c(eqn_1, "\n", eqn_2)

heights %>% 
  ggplot(aes(mother_height, daughter_height)) + 
  geom_blank() +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point(data = heights_mean, color = "red", size = 2) +
  annotate("text", label = eqns, x = 63, y = 63, hjust = 0) +
  scale_x_continuous(
    breaks = c(55, 60, heights_mean$mother_height, 65, 70),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = c(55, 60, heights_mean$daughter_height, 65, 70),
    minor_breaks = NULL
  ) +
  coord_fixed() +
  labs(
    title = "Fitted regression line and average of data",
    x = "Mother's height (inches)",
    y = "Adult daughter's height (inches)"
  )
```

Fitted regression line with x = 0.

```{r, fig.asp=1}
heights %>% 
  ggplot(aes(mother_height, daughter_height)) + 
  geom_blank() +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point(data = heights_mean, color = "red", size = 2) +
  annotate("text", label = eqn_2, x = 35, y = 45, hjust = 0) +
  scale_x_continuous(breaks = 0, minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, intercept), minor_breaks = NULL) +
  coord_fixed(xlim = c(0, NA), ylim = c(0, NA)) +
  labs(
    title = "Fitted regression line with x = 0",
    x = "Mother's height (inches)",
    y = "Adult daughter's height (inches)"
  )
```

Fitted regression line and data with x = 0.

```{r, fig.asp=1}
heights %>% 
  ggplot(aes(mother_height, daughter_height)) + 
  geom_count() +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point(data = heights_mean, color = "red", size = 2) +
  annotate("text", label = eqn_2, x = 35, y = 45, hjust = 0) +
  scale_x_continuous(
    breaks = c(0, heights_mean$mother_height),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    breaks = c(0, intercept, heights_mean$daughter_height),
    minor_breaks = NULL
  ) +
  scale_size(range = c(1, 6) / 4) +
  coord_fixed(xlim = c(0, NA), ylim = c(0, NA)) +
  labs(
    title = "Fitted regression line and data with x = 0",
    x = "Mother's height (inches)",
    y = "Adult daughter's height (inches)"
  )
```

