---
title: "Regression and Other Stories: Arsenic"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Building a logistic regression model: wells in Bangladesh. See
Chapters 13 and 14 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(haven)
library(rstanarm)

# Parameters
  # Data on arsenic in wells in Bangladesh
file_wells_all <- here::here("Arsenic/data/all.dta")
  # Data on arsenic in unsafe wells in Bangladesh
file_wells <- here::here("Arsenic/data/wells.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 13 Logistic regression

## 13.7 Building a logistic regression model: wells in Bangladesh

### Background

#### Data on all wells

Data on arsenic level in all wells and their location.

```{r}
wells_all <- 
  file_wells_all %>% 
  read_dta() %>% 
  transmute(arsenic = as / 100, x, y)

summary(wells_all)
```

The position data appear to have some problems.

```{r}
sort(wells_all$x) %>% 
  head(100)

sort(wells_all$y) %>% 
  head(100)
```

Well remove the problematic rows and center and scale the positions.

```{r}
wells_all <- 
  wells_all %>% 
  drop_na(arsenic) %>% 
  filter(x > 91, y > 24) %>% 
  mutate(
    x = (x - mean(x)) / 1000,
    y = (y - mean(y)) / 1000
  )
```

Distribution of arsenic level in all wells.

```{r}
wells_all %>% 
  ggplot(aes(arsenic)) +
  geom_histogram(binwidth = 0.1, boundary = 0) +
  geom_vline(xintercept = 0.5, color = "red") +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = "Distribution of arsenic level in all wells",
    subtitle = "Levels lower than the vertical line are considered safe",
    x = "Arsenic level",
    y = "Count"
  )
```

```{r}
quantile(wells_all$arsenic, probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1))
```

There are some outliers with levels up to 9.65, but the majority of the wells have levels under 3.

```{r}
v <- mean(wells_all$arsenic > 0.5)
v
```

About `r format(100*v, digits = 1, nsmall = 1)`% of wells are unsafe; that is, have arsenic levels over 0.5.

Map of wells in an area of Araihazar, Bangladesh.

```{r, fig.asp=1.25}
wells_all %>% 
  mutate(
    type = if_else(arsenic > 0.5, "Unsafe wells", "Safe wells"),
    arsenic = if_else(arsenic > 3, 3, arsenic)
  ) %>% 
  ggplot(aes(x, y, color = arsenic)) +
  geom_point(alpha = 0.75, size = 0.1) +
  coord_fixed() +
  facet_grid(rows = vars(type)) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  scale_y_continuous(breaks = scales::breaks_width(1)) +
  scale_color_viridis_c(breaks = c(0.5, 1:3), labels = c(c(0.5, 1:2), "3+")) +
  labs(
    title = "Map of wells in an area of Araihazar, Bangladesh",
    x = NULL,
    y = NULL,
    color = "Arsenic\nlevel"
  )
```

It looks as though the wells may be near dwellings along streets. In many areas unsafe wells are close to safe wells.

#### Data on unsafe wells

Data

```{r, message=FALSE}
wells <- read_csv(file_wells)

summary(wells)
```

The variables are:

* `switch`: Outcome variable:
  * 1 if household switched to a new well
  * 0 if household continued using its own well
* `arsenic`: Arsenic level of respondent's well
* `dist`: Distance (in meters) to the closest known safe well
* `dist100` = `dist / 100`
* `assoc`: Whether any members of the household are active in community organizations
* `educ`: Education level of the head of household
* `educ4` = `educ / 4`

##### 1D EDA

Whether household switched.

```{r}
wells %>% 
  count(switch) %>% 
  mutate(prop = n / sum(n))
```

About 57.5% of households switched.

Distribution of arsenic level in unsafe wells.

```{r}
wells %>% 
  ggplot(aes(arsenic)) +
  geom_histogram(binwidth = 0.1, boundary = 0) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = "Distribution of arsenic level in unsafe wells",
    x = "Arsenic level",
    y = "Count"
  )
```

Distance to the closest known safe well.

```{r}
wells %>% 
  ggplot(aes(dist)) +
  geom_histogram(binwidth = 10, boundary = 0) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Distance to the closest known safe well",
    x = "Distance (meters)",
    y = "Count"
  )
```

```{r}
quantile(wells$dist, probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1))
```

The median distance to a safe well is about 37 meters. About 90% of the households have a safe well within 100 meters.

##### 2D EDA

Percentage of households who switched to new well by arsenic level.

```{r, message=FALSE, fig.asp=0.75}
wells %>% 
  ggplot(aes(arsenic, switch)) +
  stat_ydensity(
    aes(group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_smooth() +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.1),
    minor_breaks = NULL,
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = 
      "Percentage of households who switched to new well by arsenic level",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Arsenic level",
    y = "Percentage of households who switched"
  )
```

As expected, the percentage of households increases with the arsenic level in their well, from about 40% for wells that are just over the safety threshold to perhaps 80% for very high levels. The sparse data for high arsenic levels results in a large uncertainty.

Percentage of households who switched to new well by distance.

```{r, message=FALSE, fig.asp=0.75}
wells %>% 
  ggplot(aes(dist, switch)) +
  stat_ydensity(
    aes(group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_smooth() +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.1),
    minor_breaks = NULL,
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Percentage of households who switched to new well by distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Percentage of households who switched"
  )
```

As expected, the percentage of households decreases with the distance to the closest known safe well, from about 60% when a safe well is very close to perhaps 20% when a safe well is far away. The sparse data for large distances results in a large uncertainty.

### Logistic regression with just one predictor

Fit a model using distance to the nearest safe well.

```{r}
set.seed(733)

fit_1 <- 
  stan_glm(
    switch ~ dist,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_1, digits = 3)
```

LOO log score

```{r}
loo_1 <- loo(fit_1)

loo_1
```

Fit a model using scaled distance to the nearest safe well.

```{r}
set.seed(733)

fit_2 <- 
  stan_glm(
    switch ~ dist100,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

fit_2
```

LOO log score

```{r}
loo_2 <- loo(fit_2)

loo_2
```

### Graphing the fitted model

Probability of household switching to new well by distance.

```{r, fig.asp=0.75}
v <- 
  tibble(
    dist = seq_range(wells$dist),
    .pred = predict(fit_1, type = "response", newdata = tibble(dist))
  )

wells %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred), data = v) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Probability of household switching to new well by distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching"
  )
```

The empirical plot in the 2D EDA section showed that about 60% of households switched at the closest distance and about 20% switched at the farthest distance. The probabilities of the model at the extremes are similar. Between the distance of 0 - 50 meters, the empirical plot has a roughly constant percentage of households who switched, whereas the model has a steady decline.

### Interpreting the logistic regression coefficients

Proportion of households who switched.

```{r}
switch_prop <- mean(wells$switch)

switch_prop
```

Log score for model with a constant prediction equal to the proportion of households who switched.

```{r}
log(switch_prop) * sum(wells$switch) + log(1 - switch_prop) * sum(!wells$switch)
```

Since the LOO log score for the model with distance is -2040, distance supplies some predictive information.

### Adding a second input variable

Fit a model using scaled distance and arsenic level.

```{r}
set.seed(733)

fit_3 <- 
  stan_glm(
    switch ~ dist100 + arsenic,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

fit_3
```

LOO log score

```{r}
loo_3 <- loo(fit_3)

loo_3
```

Compare log scores.

```{r}
loo_compare(loo_2, loo_3)
```

Model 3 has the better log score, indicating that including arsenic level in the model clearly improves the predictive accuracy.

### Graphing the fitted model with two predictors

Probability of household switching to new well by distance and arsenic level.

```{r, fig.asp=0.75}
v <- 
  tibble(
    arsenic = c(0.5, quantile(wells$arsenic, probs = c(0.25, 0.5, 0.75))),
    label = 
      case_when(
        names(arsenic) == "" ~ as.character(arsenic),
        TRUE ~ 
          str_glue(
            "{format(arsenic, digits = 1, nsmall = 1)} ({names(arsenic)})"
          ) %>% 
          as.character()
      ) %>% 
      fct_inorder(),
    dist = list(seq_range(wells$dist))
  ) %>% 
  unnest(dist) %>% 
  mutate(
    dist100 = dist / 100,
    .pred = 
      predict(fit_3, type = "response", newdata = tibble(arsenic, dist100))
  )

v %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred, color = label)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Probability of household switching to new well by distance and arsenic level",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching",
    color = "Arsenic level (Quantile)"
  )
```

The probability decreases with distance and increases with arsenic level.

Probability of household switching to new well by arsenic level and distance.

```{r, fig.asp=0.75}
v <- 
  tibble(
    dist = c(0, quantile(wells$dist, probs = c(0.25, 0.5, 0.75))),
    label = 
      case_when(
        names(dist) == "" ~ as.character(dist),
        TRUE ~ 
          str_glue(
            "{format(dist, digits = 0, nsmall = 0)} ({names(dist)})"
          ) %>% 
          as.character()
      ) %>% 
      fct_inorder(),
    arsenic = list(seq_range(wells$arsenic))
  ) %>% 
  unnest(arsenic) %>% 
  mutate(
    dist100 = dist / 100,
    .pred = 
      predict(fit_3, type = "response", newdata = tibble(arsenic, dist100))
  )

v %>% 
  ggplot(aes(arsenic)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred, color = label)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Probability of household switching to new well by arsenic level and distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Arsenic level",
    y = "Probability of household switching",
    color = "Distance in meters (Quantile)"
  )
```

The probability increases with arsenic level and decreases with distance.

# 14 Working with logistic regression

## 14.2 Logistic regression with interactions

Fit a model using scaled distance, arsenic level, and an interaction

```{r}
set.seed(733)

fit_4 <- 
  stan_glm(
    switch ~ dist100 + arsenic + dist100:arsenic,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_4, digits = 2)
```

LOO log score

```{r}
loo_4 <- loo(fit_4)

loo_4
```

### Centering the input variables

```{r}
wells <- 
  wells %>% 
  mutate(
    arsenic_c = arsenic - mean(arsenic),
    dist100_c = dist100 - mean(dist100)
  )
```

### Re-fitting the interaction model using the centered inputs

```{r}
set.seed(733)

fit_5 <- 
  stan_glm(
    switch ~ dist100_c + arsenic_c + dist100_c:arsenic_c,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_5, digits = 2)
```

LOO log score

```{r}
loo_5 <- loo(fit_5)

loo_5
```

Compare log scores.

```{r}
loo_compare(loo_4, loo_5)
```

Centering the variables does not affect the model log score.

### Statistical significance of the interaction

Compare log scores.

```{r}
loo_compare(loo_3, loo_4)
```

Adding the interaction doesn't change the predictive performance, and there is no need to keep it in the model for predictive purposes (unless new information can be obtained).

### Graphing the model with interactions

Probability of household switching to new well by distance and arsenic level.

```{r, fig.asp=0.75}
v <- 
  tibble(
    arsenic = c(0.5, quantile(wells$arsenic, probs = c(0.25, 0.5, 0.75))),
    label = 
      case_when(
        names(arsenic) == "" ~ as.character(arsenic),
        TRUE ~ 
          str_glue(
            "{format(arsenic, digits = 1, nsmall = 1)} ({names(arsenic)})"
          ) %>% 
          as.character()
      ) %>% 
      fct_inorder(),
    dist = list(seq_range(wells$dist))
  ) %>% 
  unnest(dist) %>% 
  mutate(
    dist100 = dist / 100,
    .pred = 
      predict(fit_4, type = "response", newdata = tibble(arsenic, dist100))
  )

v %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred, color = label)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Probability of household switching to new well by distance and arsenic level",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching",
    color = "Arsenic level (Quantile)"
  )
```

The probability decreases with distance and increases with arsenic level.

Probability of household switching to new well by arsenic level and distance.

```{r, fig.asp=0.75}
v <- 
  tibble(
    dist = c(0, quantile(wells$dist, probs = c(0.25, 0.5, 0.75))),
    label = 
      case_when(
        names(dist) == "" ~ as.character(dist),
        TRUE ~ 
          str_glue(
            "{format(dist, digits = 0, nsmall = 0)} ({names(dist)})"
          ) %>% 
          as.character()
      ) %>% 
      fct_inorder(),
    arsenic = list(seq_range(wells$arsenic))
  ) %>% 
  unnest(arsenic) %>% 
  mutate(
    dist100 = dist / 100,
    .pred = 
      predict(fit_4, type = "response", newdata = tibble(arsenic, dist100))
  )

v %>% 
  ggplot(aes(arsenic)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred, color = label)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Probability of household switching to new well by arsenic level and distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Arsenic level",
    y = "Probability of household switching",
    color = "Distance in meters (Quantile)"
  )
```

The probability increases with arsenic level and decreases with distance.

### Adding social predictors

Fit a model using scaled distance, arsenic level, education of head of household, and community organization activity.

```{r}
set.seed(733)

fit_6 <- 
  stan_glm(
    switch ~ dist100 + arsenic + educ4 + assoc,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_6, digits = 2)
```

LOO log score

```{r}
loo_6 <- loo(fit_6)

loo_6
```

Compare log scores.

```{r}
loo_compare(loo_4, loo_6)
```

Belonging to a community association, perhaps surprisingly, is associated in our data with a _lower_ probability of switching, after adjusting for the other factors in the model. However, this coefficient is not estimated precisely, and so for clarity and stability we remove it from the model.

Fit a model using scaled distance, arsenic level, and education of head of household.

```{r}
set.seed(733)

fit_7 <- 
  stan_glm(
    switch ~ dist100 + arsenic + educ4,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_7, digits = 2)
```

LOO log score

```{r}
loo_7 <- loo(fit_7)

loo_7
```

Compare log scores.

```{r}
loo_compare(loo_4, loo_7)
```

Adding education improves predictive log score, but there is considerable uncertainty.

```{r}
loo_compare(loo_6, loo_7)
```

Removing the association variable doesn't change the predictive performance.

### Adding further interactions

Create centered education variable.

```{r}
wells <- 
  wells %>% 
  mutate(educ4_c = educ4 - mean(educ4))
```

Fit a model using scaled distance, arsenic level, education of head of household, and interactions with education.

```{r}
set.seed(733)

fit_8 <- 
  stan_glm(
    switch ~ 
      dist100_c + arsenic_c + educ4_c + dist100_c:educ4_c + arsenic_c:educ4_c,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_8, digits = 2)
```

LOO log score

```{r}
loo_8 <- loo(fit_8)

loo_8
```

Compare log scores.

```{r}
loo_compare(loo_7, loo_8)
```

Adding the interactions with education in model 8 improves predictive performance over model 7.

```{r}
loo_compare(loo_3, loo_8)
```

The education variable and its interactions in model 8 substantially improves predictive performance over model 3.

## 14.3 Predictive simulation

### Simulating the uncertainty in the estimated coefficients

Posterior draws of logistic regression coefficients.

```{r}
sims_2 <- as_tibble(fit_2)
coef <- 
  tibble(
    `(Intercept)` = coef(fit_2)[["(Intercept)"]],
    dist100 = coef(fit_2)[["dist100"]]
  )

sims_2 %>% 
  ggplot(aes(`(Intercept)`, dist100)) +
  geom_point(size = 0.1) +
  geom_point(data = coef, color = "red", size = 1.5) +
  labs(
    title = "Posterior draws of logistic regression coefficients"
  )
```

Probability of household switching to new well by distance with uncertainty: With 50% and 90% predictive intervals.

```{r, fig.asp=0.75}
new <- tibble(dist = seq_range(wells$dist))
linpred <- posterior_linpred(fit_1, newdata = new)
v <- 
  new %>% 
  mutate(
    .pred = predict(fit_1, type = "response", newdata = new),
    `5%`  = apply(linpred, 2, quantile, probs = 0.05) %>% plogis(),
    `25%` = apply(linpred, 2, quantile, probs = 0.25) %>% plogis(),
    `75%` = apply(linpred, 2, quantile, probs = 0.75) %>% plogis(),
    `95%` = apply(linpred, 2, quantile, probs = 0.95) %>% plogis()
  )

v %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), alpha = 0.25) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), alpha = 0.5) +
  geom_line(aes(y = .pred)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Probability of household switching to new well by distance",
    subtitle =
      "With 50% and 90% predictive intervals\nVoilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching"
  )
```

In the region of sparse data for large distances, the uncertainty from the posterior distribution of the model is much less than the uncertainty seen in the [2D EDA section](#2d-eda) using LOESS.

