---
title: "Regression and Other Stories: Arsenic"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Average predictive comparisons for a logistic regression model:
wells in Bangladesh. See Chapter 14 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Data on arsenic in unsafe wells in Bangladesh
file_wells <- here::here("Arsenic/data/wells.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 14 Working with logistic regression

## 14.4 Average predictive comparisons on the probability scale

### Demonstration with the well-switching example

Data

```{r, message=FALSE}
wells <- read_csv(file_wells)

summary(wells)
```

The variables are:

* `switch`: Outcome variable:
  * 1 if household switched to a new well
  * 0 if household continued using its own well
* `arsenic`: Arsenic level of respondent's well
* `dist`: Distance (in meters) to the closest known safe well
* `dist100` = `dist / 100`
* `assoc`: Whether any members of the household are active in community organizations
* `educ`: Education level of the head of household
* `educ4` = `educ / 4`

Fit a model using scaled distance, arsenic level, and education of head of household.

```{r}
set.seed(733)

fit_7 <- 
  stan_glm(
    switch ~ dist100 + arsenic + educ4,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_7, digits = 2)
```

#### Average predictive difference in probability of switching

Model coefficients.

```{r}
b <- coef(fit_7)
names(b)
```

Compare households that are next to or 100 meters from the nearest safe well.

```{r}
dist100_lo <- 0
dist100_hi <- 1

dist100_apd <- 
  wells %>% 
  mutate(
    delta = 
      plogis(b[1] + b[2] * dist100_hi + b[3] * arsenic + b[4] * educ4) -
      plogis(b[1] + b[2] * dist100_lo + b[3] * arsenic + b[4] * educ4)
  ) %>% 
  pull(delta) %>% 
  mean()

dist100_apd
```

The result is `r format(dist100_apd, digits = 2, nsmall = 2)`, implying that, on average in the data, households that are 100 meters from the nearest safe well are `r round(100 * abs(dist100_apd))`% less likely to switch, compared to households that are right next to the nearest safe well, at the same arsenic and education levels.

#### Comparing probabilities of switching for households differing in arsenic levels

Compare households with existing arsenic levels of 0.5 or 1.0.

```{r}
arsenic_lo <- 0.5
arsenic_hi <- 1.0

arsenic_apd <- 
  wells %>% 
  mutate(
    delta = 
      plogis(b[1] + b[2] * dist100 + b[3] * arsenic_hi + b[4] * educ4) -
      plogis(b[1] + b[2] * dist100 + b[3] * arsenic_lo + b[4] * educ4)
  ) %>% 
  pull(delta) %>% 
  mean()

arsenic_apd
```

The result is `r format(arsenic_apd, digits = 2, nsmall = 3)` -- so this corresponds to a `r format(100 * abs(arsenic_apd), digits = 1, nsmall = 1)`% difference in the probability of switching.

#### Average predictive difference in probability of switching, comparing households with 0 and 12 years of education

Compare households with 0 or 12 years of education.

```{r}
educ4_lo <- 0
educ4_hi <- 3

educ4_apd <- 
  wells %>% 
  mutate(
    delta = 
      plogis(b[1] + b[2] * dist100 + b[3] * arsenic + b[4] * educ4_hi) -
      plogis(b[1] + b[2] * dist100 + b[3] * arsenic + b[4] * educ4_lo)
  ) %>% 
  pull(delta) %>% 
  mean()

educ4_apd
```

This comes to `r format(educ4_apd, digits = 2, nsmall = 2)`, a difference of `r round(100 * abs(educ4_apd))`%.

### Average predictive comparisons in the presence of interactions

Center variables.

```{r}
wells <- 
  wells %>% 
  mutate(
    arsenic_c = arsenic - mean(arsenic),
    dist100_c = dist100 - mean(dist100),
    educ4_c = educ4 - mean(educ4)
  )
```

Fit a model using scaled distance, arsenic level, education of head of household, and interactions with education.

```{r}
set.seed(733)

fit_8 <- 
  stan_glm(
    switch ~ 
      dist100_c + arsenic_c + educ4_c + dist100_c:educ4_c + arsenic_c:educ4_c,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_8, digits = 2)
```

Model coefficients.

```{r}
b <- coef(fit_8)
names(b)
```

Compare households that are next to or 100 meters from the nearest safe well.

```{r}
dist100_lo <- 0
dist100_hi <- 1
dist100_c_lo <- dist100_lo - mean(wells$dist100)
dist100_c_hi <- dist100_hi - mean(wells$dist100)

dist100_apd <- 
  wells %>% 
  mutate(
    delta = 
      plogis(
        b[1] + b[2] * dist100_c_hi + b[3] * arsenic_c + b[4] * educ4_c +
          b[5] * dist100_c_hi * educ4_c + b[6] * arsenic_c * educ4_c
      ) -
      plogis(
        b[1] + b[2] * dist100_c_lo + b[3] * arsenic_c + b[4] * educ4_c +
          b[5] * dist100_c_lo * educ4_c + b[6] * arsenic_c * educ4_c
      )
  ) %>% 
  pull(delta) %>% 
  mean()

dist100_apd
```

This comes to `r format(dist100_apd, digits = 2, nsmall = 2)`, a difference of `r round(100 * abs(dist100_apd))`%.

