---
title: "Regression and Other Stories: National election study"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Fitting the same regression to many datasets. See Chapter 10 in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # National Election Study data
file_nes <- here::here("NES/data/nes.txt")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 10 Linear regression with multiple predictors

## 10.9 Fitting the same model to many datasets

### Predicting party identification

Data

```{r}
nes <-
  file_nes %>% 
  read.table() %>% 
  as_tibble()

glimpse(nes)
```

Note that the data has weight variables `weight*`. It may be appropriate to perform a weighted linear regression. But since we have no documentation for these variables, we will ignore them.

Our models will use the following variables for the years 1976 - 2000.

```{r}
nes <- 
  nes %>% 
  filter(year >= 1976, year <= 2000) %>% 
  select(
    year,
    partyid7,
    real_ideo,
    race_adj,
    age_discrete,
    educ1,
    female,
    income
  )

summary(nes)
```

The variable `real_ideo` has `r format(100 * mean(is.na(nes$real_ideo)), digits = 1, nsmall = 1)`% `NA`s. This could be a problem, but we will ignore it.

Finally, we'll convert `age_discrete` into a factor.

```{r}
nes <- 
  nes %>% 
  mutate(
    age =
      case_when(
        age_discrete == 1 ~ "18 - 29",
        age_discrete == 2 ~ "30 - 44",
        age_discrete == 3 ~ "45 - 64",
        age_discrete == 4 ~ "65+",
        TRUE ~ NA_character_
      ) %>% 
      factor()
  ) %>% 
  select(!age_discrete)
```

Fit the a linear regression for each election year. Return the coefficients and their standard errors for each regression.

```{r}
set.seed(178)

coefs <- 
  nes %>% 
  nest(data = !year) %>% 
  mutate(
    fit =
      map(
        data,
        ~ stan_glm(
          partyid7 ~ real_ideo + race_adj + age + educ1 + female + income,
          data = .,
          refresh = 0
        )
      ),
    coefs =
      map(
        fit,
        ~ left_join(
          enframe(coef(.), name = "var", value = "coef"),
          enframe(se(.), name = "var", value = "se"),
          by = "var"
        )
      )
  ) %>% 
  select(!c(data, fit)) %>% 
  unnest(cols = coefs)

coefs
```

Linear regression coefficients by election year: With 50% uncertainty intervals.

```{r, fig.asp=0.75}
var_labels <- 
  c(
    "(Intercept)" = "Intercept",
    "real_ideo" = "Ideology",
    "race_adj" = "Black",
    "age30 - 44" = "Age 30 - 44",
    "age45 - 64" = "Age 45 - 64",
    "age65+" = "Age 65+",
    "educ1" = "Education",
    "female" = "Female",
    "income" = "Income"
  )

coefs %>% 
  mutate(
    var = fct_inorder(var),
    q_25 = qnorm(0.25, mean = coef, sd = se),
    q_75 = qnorm(0.75, mean = coef, sd = se)
  ) %>% 
  ggplot(aes(year, coef)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_line() +
  geom_linerange(aes(ymin = q_25, ymax = q_75)) +
  geom_point() +
  facet_wrap(
    vars(var),
    ncol = 3,
    scales = "free_y",
    labeller = labeller(var = var_labels)
  ) +
  scale_x_continuous(breaks = seq(1976, 2000, 4)) +
  theme(axis.text = element_text(size = rel(0.6))) +
  labs(
    title = "Linear regression coefficients by election year",
    subtitle = "With 50% uncertainty intervals",
    x = "Election year",
    y = "Coefficient"
  )
```

