---
title: "Regression and Other Stories: ChildCare"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

 Code and figures for Infant Health and Development Program (IHDP)
 example. See Chapter 20 in Regression and Other Stories.

The intervention for low-birth-weight children is described by

- Brooks-Gunn, J., Liaw, F. R., and Klebanov, P. K. (1992). Effects
  of early intervention on cognitive function of low birth weight
  preterm infants. Journal of Pediatrics 120, 350–359.
- Hill, J. L., Brooks-Gunn, J., and Waldfogel, J. (2003). Sustained
  effects of high participation in an early intervention for
  low-birth-weight premature infants. Developmental Psychology 39,
  730–744.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Infant Health and Development Program (IHDP) data
file_cc2 <- here::here("Childcare/data/cc2.csv") 
  # Variables used in analysis
vars <- 
  vars(
    momage, b_marr, work_dur, prenatal, sex, first, bw, preterm, age, dayskidh,
    ppvtr_36, st5, st9, st12, st25, st36, st42, st48, st53, income, treat,
    black, hispanic, white, lths, hs, ltcoll, college, bwg, educ
  )
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 20 Observational studies with measured confounders

## 20.5 Example: evaluating a child care program

Data

```{r, message=FALSE}
cc2 <- 
  file_cc2 %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  select(!!! vars)

glimpse(cc2)
```

The variables are:

* `momage`: Mother's age at time of birth.
* `b_marr`: Indicator for whether mother was married at birth.
* `work_dur`: Indicator for whether mother worked in the year before she gave birth.
* `prenatal`: Indicator for whether mother received prenatal care.
* `sex`: Indicator for whether child was born male or female.
* `first`: Indicator for whether child was the first born for the mother.
* `bw`: Child's birth weight (grams).
* `preterm`: Number of weeks preterm child was born.
* `age`: Age of child (months).
* `dayskidh`: Number of days child was in the hospital after being born.
* `ppvtr_36`: IQ measured at age 36 months.
* `st5`, `st9`, `st12`, `st25`, `st36`, `st42`, `st48`, `st53`: Indicator for state where household resides.
* `income`: Family income one year after the child was born.
* `treat`: Indicator for whether family was allowed to receive IHDP services (1 = yes).
* `black`, `hispanic`, `white`: Indicators for child’s race/ethnicity.
* `lths`, `hs`, `ltcoll`, `college`: Indicators for mother’s education at time of birth.
* `bwg`: Indicator for whether child was born low birth weight.
* `educ`: Mother's education at time of birth.

```{r}
summary(cc2)
```

We can see from the above that there are no `NA`s in the data.

```{r}
v <- 
  cc2 %>% 
  count(negative = income < 0)

v
```

`income` has `r v$n[v$negative]` negative values.

```{r}
cc2 %>% 
  select(!c(momage, bw, preterm, age, dayskidh, ppvtr_36, income, educ)) %>% 
  unlist() %>% 
  setequal(0:1)
```

And all variables other than `momage`, `bw`, `preterm`, `age`, `dayskidh`, `ppvtr_36`, `income`, and `educ` are indicator variables.

```{r}
cc2 %>% 
  count(educ, lths, hs, ltcoll, college)
```

`educ` combines the indicators variables `lths`, `hs`, `ltcoll`, and `college`.

### Examining imbalances for several covariates

Children in Infant Health and Development Program.

```{r, fig.asp=0.75}
fit <- stan_glm(ppvtr_36 ~ bw + treat, data = cc2, refresh = 0)

lines <- 
  tribble(
      ~treat, ~intercept, ~slope,
      0, coef(fit)[["(Intercept)"]], coef(fit)[["bw"]],
      1, coef(fit)[["(Intercept)"]] + coef(fit)[["treat"]], coef(fit)[["bw"]]
    ) %>% 
  mutate(treat = factor(treat, labels = c("Control", "Treatment")))

cc2 %>% 
  mutate(treat = factor(treat, labels = c("Control", "Treatment"))) %>% 
  arrange(treat) %>% 
  ggplot(aes(bw, ppvtr_36, color = treat)) +
  geom_point(size = 0.5, alpha = 0.75) +
  geom_abline(
    aes(slope = slope, intercept = intercept, color = treat),
    data = lines
  ) +
  coord_cartesian(xlim = c(NA, 5000)) +
  scale_y_continuous(breaks = scales::breaks_width(20)) +
  scale_color_discrete(direction = -1) +
  theme(legend.position = "bottom") +
  labs(
    title = "Children in Infant Health and Development Program",
    x = "Birth weight",
    y = "Test score at age 3",
    color = "Group"
  )
```

Distribution of mother's education.

```{r}
educ_labels <- 
  c(
    "Not a high school\ngraduate",
    "High school\ngraduate",
    "Some college",
    "College graduate"
  )

cc2 %>% 
  mutate(
    educ = factor(educ, labels = educ_labels),
    treat = factor(treat, labels = c("Control", "Treatment"))
  ) %>% 
  count(educ, treat) %>% 
  group_by(treat) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup() %>% 
  ggplot(aes(educ, prop, fill = treat)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_fill_discrete(direction = -1) +
  labs(
    title = "Distribution of mother's education",
    x = "Mother's education at time of birth",
    y = "Percentage of group",
    fill = "Group"
  )
```

Each category includes observations in each treatment group. However, the percentages falling in each category differ when comparing treatment and control groups -- thus there is clearly an imbalance.

Distribution of children's ages.

```{r}
cc2 %>% 
  mutate(treat = factor(treat, labels = c("Control", "Treatment"))) %>% 
  ggplot(aes(age, y = stat(density))) +
  geom_histogram(binwidth = 20, boundary = 0) +
  facet_grid(rows = vars(treat)) +
  scale_x_continuous(breaks = scales::breaks_width(20)) +
  labs(
    title = "Distribution of children's ages",
    x = "Age of child (months)",
    y = "Density"
  )
```

The averages of children's ages differ little across treatment groups, but the vast majority of control children have ages that are not represented in the treatment group.

## 20.6 Subclassification and average treatment effects

### Subclassification as a simple nonparameteric approach

Difference in mean test scores within each subclass defined by mother's education, as well as the associated standard error and sample size in each subclass.

```{r}
v <- 
  cc2 %>% 
  group_by(educ) %>% 
  summarize(
    treatment_effect = mean(ppvtr_36[treat == 1]) - mean(ppvtr_36[treat == 0]),
    var_treatment = var(ppvtr_36[treat == 1]),
    n_treatment = sum(treat == 1),
    var_control = var(ppvtr_36[treat == 0]),
    n_control = sum(treat == 0),
    std_error = sqrt(var_treatment / n_treatment + var_control / n_control)
  ) %>% 
  select(educ, treatment_effect, std_error, n_treatment, n_control)


v %>% 
  mutate(
    educ = case_when(
      educ == 1 ~ "Not a high school graduate",
      educ == 2 ~ "High school graduate",
      educ == 3 ~ "Some college",
      educ == 4 ~ "College graduate",
      TRUE ~ NA_character_
    )
  ) %>% 
  knitr::kable()
```

Estimated average treatment effect.

```{r}
v <- 
  v %>% 
  mutate(n = n_treatment + n_control)

weighted.mean(v$treatment_effect, v$n)
```

Standard error.

```{r}
sqrt(sum(v$std_error^2 * v$n^2) / sum(v$n)^2)
```

Average effect of the treatment on the treated.

```{r}
weighted.mean(v$treatment_effect, v$n_treatment)
```

Standard error.

```{r}
sqrt(sum(v$std_error^2 * v$n_treatment^2) / sum(v$n_treatment)^2)
```

---

The remainder of this example uses functions defined [here](https://github.com/avehtari/ROS-Examples/tree/master/Childcare/library) and from the survey package. The code for calling these functions can be found [here](https://avehtari.github.io/ROS-Examples/Childcare/childcare.html).

