---
title: "Regression and Other Stories: Sesame street"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Causal analysis of Sesame Street experiment. See Chapters 18 and 21
in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(brms)
library(rstanarm)

# Parameters
  # Seed
SEED <- 1234
  # Results of educational experiment
file_sesame <- here::here("Sesame/data/sesame.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 18 Causal inference and randomized experiments

## 18.8 Exercises

Data

```{r, message=FALSE}
sesame <- 
  file_sesame %>% 
  read_csv() %>% 
  mutate(
    site =
      case_when(
        `_Isite_2` == 1 ~ 2,
        `_Isite_3` == 1 ~ 3,
        `_Isite_4` == 1 ~ 4,
        `_Isite_5` == 1 ~ 5,
        TRUE ~ 1
      ) %>% 
      factor()
  ) %>% 
  relocate(site, .before = `_Isite_2`)

glimpse(sesame)
```

Let's check the derived variable `site`.

```{r}
sesame %>% 
  count(site, across(starts_with("_Isite")))
```


# 21 Additional topics in causal inference

## 21.1 Estimating causal effects indirectly using instrumental variables

### Instrumental variables estimate: Sesame Street

Let's first look at the effect of encouraging children to watch Sesame Street.

```{r}
v <- 
  sesame %>% 
  count(encouraged, watched) %>% 
  group_by(encouraged) %>% 
  mutate(watched_prop = n / sum(n)) %>% 
  ungroup()

v
```

Without any encouragement, `r round(100 * (v %>% filter(encouraged == 0, watched == 1) %>% pull(watched_prop)))`% of children still watched Sesame Street. With encouragement, that percentage increased to `r round(100 * (v %>% filter(encouraged == 1, watched == 1) %>% pull(watched_prop)))`%

We now calculate an estimate of the effect of watching Sesame Street for the children who received encouragement and who in turn actually watched the show, the compliers.

We first estimate the percentage of children actually induced to watch Sesame Street by the intervention, which is the coefficient on the instrument (`encouraged`), in the following regression:

```{r}
set.seed(264)

fit_1a <- stan_glm(watched ~ encouraged, data = sesame, refresh = 0)

print(fit_1a, digits = 2)
```

The estimated coefficient of `encouraged` here is `r format(coef(fit_1a)[["encouraged"]], digits = 2, nsmall = 2)`. This is simply the difference in the proportion of children who watched Sesame Street when encouraged vs. when not encouraged:

```{r}
v %>% 
  filter(watched == 1) %>% 
  summarize(
    watched_diff = watched_prop[encouraged == 1] - watched_prop[encouraged == 0]
  )
```

We then compute the intent-to-treat estimate, obtained in this case using the regression of the outcome (the _post_-treatment measurement of the _letter_ recognition task, `postlet`) on the instrument:

```{r}
set.seed(264)

fit_1b <- stan_glm(postlet ~ encouraged, data = sesame, refresh = 0)

print(fit_1b, digits = 2)
```

The estimated coefficient of `encouraged` in this regression is `r format(coef(fit_1b)[["encouraged"]], digits = 1, nsmall = 1)`, which we then "inflate" by dividing by the percentage of children affected by the intervention:

```{r}
wald_est <- coef(fit_1b)[["encouraged"]] / coef(fit_1a)[["encouraged"]]

wald_est
```

This ratio is sometimes called the _Wald estimate_.

## 21.2 Instrumental variables in a regression framework

### Two-stage least squares: Sesame Street

The first step is to regress the "treatment" variable -- an indicator for regular watching (`watched`) -- on the randomized instrument, encouragement to watch (`encouraged`). This is the same as `fit_1a` above, but we run again here and relabel to indicate that in this framework it is our first-stage model.

```{r}
set.seed(264)

fit_2a <- stan_glm(watched ~ encouraged, data = sesame, refresh = 0)

print(fit_2a, digits = 2)
```

Add the predicted values for `watched`.

```{r}
sesame <- 
  sesame %>% 
  mutate(watched_pred_2a = predict(fit_2a))

summary(sesame$watched_pred_2a)
```

The second step is to regress the letter recognition outcome on the predicted values for `watched`.

```{r}
set.seed(264)

fit_2b <- stan_glm(postlet ~ watched_pred_2a, data = sesame, refresh = 0)

print(fit_2b, digits = 2)
```

Here the coefficient of `watched_pred_2a` is the effect of watching Sesame Street on letter recognition for those who would watch if encouraged but not otherwise (compliers). This second-stage regression does not give the correct standard error, however, as we discuss below.

### Adjusting for covariates in an instrument variables framework

The randomization for this particular experiment took place within sites and settings; it is therefore appropriate to adjust for these covariates in estimating the treatment effect. Additionally, pre-treatment scores are available that are highly predictive of post-test scores.

The first step of two-stage approach adjusting for these covariates:

```{r}
set.seed(264)

fit_3a <- 
  stan_glm(
    watched ~ encouraged + prelet + site + setting, 
    data = sesame, 
    refresh = 0
  )

print(fit_3a, digits = 2)
```

Add the predicted values for `watched`.

```{r}
sesame <- 
  sesame %>% 
  mutate(watched_pred_3a = predict(fit_3a))

summary(sesame$watched_pred_3a)
```

The second step is to regress the letter recognition outcome on the predicted values for `watched`.

```{r}
set.seed(264)

fit_3b <-
  stan_glm(
    postlet ~ watched_pred_3a + prelet + site + setting, 
    data = sesame, 
    refresh = 0
  )

print(fit_3b, digits = 2)
```

The estimated effect of watching Sesame Street on the compliers is about `r format(coef(fit_3b)[["watched_pred_3a"]], digits = 1, nsmall = 1)` points on the letter recognition test. Again, we do not trust this standard error and will discuss next how to appropriately adjust it for the two stages of estimation.

### Standard errors for instrumental variables estimates

We show here how to adjust the standard error to account for the uncertainty in both stages of the model.

```{r}
resid <- residuals(fit_3b)
x_adj <- model.matrix(fit_3b)
x_adj[, "watched_pred_3a"] <- sesame$watched
resid_adj <- sesame$postlet - x_adj %*% coef(fit_3b)
se_adj <- 
  se(fit_3b)[["watched_pred_3a"]] * sqrt(sum(resid^2) / sum(resid_adj^2))

se_adj
```

In this example, the resulting standard error of `r format(se_adj, digits = 2, nsmall = 2)` is slightly larger than the unadjusted standard error of `r format(se(fit_3b)[["watched_pred_3a"]], digits = 2, nsmall = 2)`.

### Performing two-stage least squares automatically using brms

It is also possible to perform a Bayesian version of the two-stage least squares model directly using brms. Here is how to estimate the effect of regularly watching Sesame Street on post-treatment letter recognition using encouragement as an instrument:

```{r, message=FALSE, warning=FALSE, error=FALSE, results=FALSE}
set.seed(264)

fit_brm_1 <- 
  brm(
    formula = bf(watched ~ encouraged) + bf(postlet ~ watched),
    data = sesame,
    refresh = 0,
    seed = SEED
  )
```

```{r}
fit_brm_1
```

The resulting estimate is the coefficient `postlet_watched`, which is the instrumental variables estimate of `watched` on `postlet`; the estimate is `r format(fixef(fit_brm_1)["postlet_watched", "Estimate"], digits = 1, nsmall = 1)` with a standard error of `r format(fixef(fit_brm_1)["postlet_watched", "Est.Error"], digits = 1, nsmall = 1)`.

To incorporate other pre-treatment variables as controls, we must include them in both stages of the regression model; for example,

```{r, message=FALSE, warning=FALSE, error=FALSE, results=FALSE}
set.seed(264)

fit_brm_2 <- 
  brm(
    formula = 
      bf(watched ~ encouraged + prelet + setting + site) +
      bf(postlet ~ watched + prelet + setting + site),
    data = sesame,
    refresh = 0,
    seed = SEED
  )
```

```{r}
fit_brm_2
```

The resulting estimate is the coefficient `postlet_watched`, which is the instrumental variables estimate of `watched` on `postlet`; the estimate is `r format(fixef(fit_brm_2)["postlet_watched", "Estimate"], digits = 1, nsmall = 1)` with a standard error of `r format(fixef(fit_brm_2)["postlet_watched", "Est.Error"], digits = 1, nsmall = 1)`.

