---
title: "Regression and Other Stories: Unemployment"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Time series fit and posterior predictive model checking for
unemployment series. See Chapter 11 in Regression and Other
Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(bayesplot)
library(rstanarm)

# Parameters
  # U.S. unemployment data
file_unemployment <- here::here("Unemployment/data/unemp.txt")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 11 Assumptions, diagnostics, and model evaluation

## 11.5 Example: predictive simulation to check the fit of a time-series model

Data

```{r, message=FALSE}
unemployment <- 
  file_unemployment %>% 
  read_table2() %>% 
  mutate(y_lag = lag(y))

unemployment
```

```{r}
summary(unemployment)
```

### Fitting a first-order autoregression to the unemployment series

U.S. annual unemployment rate.

```{r}
unemployment %>% 
  ggplot(aes(year, y)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = scales::breaks_width(10)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "U.S. annual unemployment rate",
    x = "Year",
    y = "Unemployment rate"
  )
```

Fit first-order autoregression to the unemployment series.

```{r}
set.seed(264)

fit <- stan_glm(y ~ y_lag, data = unemployment, refresh = 0)

print(fit, digits = 2)
```

### Simulating replicated datasets

```{r}
set.seed(457)

sims <- as_tibble(fit)

years <- seq(min(unemployment$year), max(unemployment$year))
y_1 <- unemployment$y[unemployment$year == years[1]]

n_sims <- nrow(sims)
n_years <- length(years)

unemployment_sim <- function(intercept, slope, sigma) {
  y <- double(length = n_years)
  y[1] <- y_1
  for (i in seq_len(n_years - 1)) {
    y[i + 1] = intercept + slope * y[i] + rnorm(1, mean = 0, sd = sigma)
  }
  tibble(year = years, y = y)
}

y_rep <- 
  sims %>% 
  pmap_dfr(
    ~ unemployment_sim(intercept = ..1, slope = ..2, sigma = ..3),
    .id = "rep"
  ) %>% 
  mutate(rep = as.integer(rep))

y_rep
```

`y_rep` is a tidy tibble with `r n_sims` * `r n_years` rows.

### Visual and numerical comparisons of replicated to actual data

Plot 20 simulated unemployment rate time series.

```{r, fig.asp=0.75}
set.seed(926)

y_rep %>% 
  filter(rep %in% sample(n_sims, 20)) %>% 
  ggplot(aes(year, y)) + 
  geom_line() +
  facet_wrap(vars(rep), ncol = 5) +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  labs(
    title = "Simulated U.S. annual unemployment rate",
    x = "Year",
    y = "Unemployment rate"
  )
```

Numerical posterior predictive check.

Calculate the number of years in which the direction of unemployment switches, that is, when an increase in unemployment is followed by a decrease, or vice versa.

```{r}
test <- function(y) {
  sum(sign(y - lag(y)) != sign(lag(y) - lag(y, n = 2L)), na.rm = TRUE)
}
```

Compare this statistic for the actual data to that for the replicates.

```{r}
test_y <- test(unemployment$y)

test_y

test_y_rep <- 
  y_rep %>% 
  group_by(rep) %>% 
  summarize(test = test(y))

summary(test_y_rep$test)

v <- mean(test_y_rep$test > test_y)
v
```

`r format(100 * v, digits = 1, nsmall = 1)`% of the replicates have more direction switches than the actual data.

Plot test statistic for data and replicates.

```{r}
test_y_rep %>% 
  ggplot(aes(test)) +
  geom_bar() +
  geom_vline(xintercept = test_y, color = "red") +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  labs(
    title = "Distribution of direction switches in replicates",
    subtitle = "Vertical line is number of direction switches in data",
    x = "Number of direction switches in replicate",
    y = "Count"
  )
```

Plot test statistic for data and replicates using bayesplot.

```{r}
v <- matrix(y_rep$y, nrow = n_sims, ncol = n_years, byrow = TRUE)

ppc_stat(y = unemployment$y, yrep = v, stat = test, binwidth = 1)
```

