---
title: "Regression and Other Stories: Gay"
author: "Andrew Gelman, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Simple models (linear and discretized age) and attitudes as a
function of age. See Chapter 12 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Data on support for same-sex marriage
file_data <- here::here("Gay/data/naes04.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 12 Transformations and regression

## 12.5 Other transformation

### Index and indicator variables

Data

```{r, message=FALSE, warning=FALSE}
data <- 
  file_data %>% 
  read_csv() %>% 
  drop_na(age, gayFavorStateMarriage) %>% 
  select(age, gayFavorStateMarriage)

data  
```

```{r}
summary(data$age)
```

`age` ranges from 18 to 97.

```{r}
data %>% 
  count(age) %>% 
  arrange(desc(age))
```

Because of the relatively small number of respondents over age 90, we will let age 91 represent respondents ages 91 and over. We will also create a variable for binned ages.

```{r}
age_max <-  91

data <- 
  data %>% 
  mutate(age = if_else(age >= age_max , age_max, age)) %>%
  group_by(age) %>% 
  summarize(favor = sum(gayFavorStateMarriage == "Yes") / n()) %>% 
  mutate(
    age_bin = cut(age, breaks = c(min(age) - 1, seq(29, 79, 10), age_max))
  )

data
```

Fit linear regression model.

```{r}
set.seed(235)

fit_linear <- stan_glm(favor ~ age, data = data, refresh = 0)

print(fit_linear, digits = 4)
```

Attitudes toward same-sex marriage by age in 2004.

```{r}
intercept <- coef(fit_linear)[["(Intercept)"]]
slope <- coef(fit_linear)[["age"]]

data %>% 
  ggplot() +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point(aes(age, favor)) +
  coord_cartesian(ylim = c(0, NA)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(
    title = "Attitudes toward same-sex marriage by age in 2004",
    x = "Age",
    y = "Favor same-sex marriage"
  )
```

Fit model with binned ages.

```{r}
set.seed(235)

fit_binned <- stan_glm(favor ~ age_bin, data = data, refresh = 0)

print(fit_binned, digits = 2)
```

Attitudes toward same-sex marriage by age in 2004: For binned ages.

```{r}
levels <- 
  coef(fit_binned) %>% 
  enframe(value = "favor") %>% 
  mutate(name = if_else(name == "(Intercept)", "17,29", name)) %>% 
  extract(
    col = name,
    into = c("age_min", "age_max"),
    regex = "(\\d+),(\\d+)",
    convert = TRUE
  ) %>% 
  mutate(
    age_min = age_min + 1L,
    favor = 
      case_when(
        age_min == 18 ~ favor,
        TRUE ~ favor[age_min == 18] + favor
      )
  )

data %>% 
  ggplot() +
  geom_segment(
    aes(x = age_min, xend = age_max, y = favor, yend = favor), 
    data = levels
  ) +
  geom_point(aes(age, favor)) +
  coord_cartesian(ylim = c(0, NA)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(
    title = "Attitudes toward same-sex marriage by age in 2004",
    subtitle = "For binned ages",
    x = "Age",
    y = "Favor same-sex marriage"
  )
```

