---
title: "Regression and Other Stories: Last letters of names"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r format(Sys.Date())`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Last letters of names - Distributions of last letters of names of
American babies. See Chapter 2 in Regression and Other Stories.

-------------

# 2 Data and measurement

## 2.3 All graphs are comparisons

### Multiple plots

```{r, message=FALSE}
# Packages
library(tidyverse)
library(babynames)

# Parameters
  # Common code
file_common <- here::here("_common.R")
  
#===============================================================================

# Run common code
source(file_common)
```

Data

```{r}
babynames
```


Last letter of boys' names by year.

```{r}
boys_last_letter <- 
  babynames %>% 
  filter(sex == "M") %>% 
  mutate(last_letter = str_sub(name, start = -1L)) %>% 
  group_by(year, last_letter) %>%
  summarize(prop = sum(prop)) %>% 
  ungroup()

boys_last_letter
```

Last letter of boys' names by year.

```{r, fig.asp=1}
boys_last_letter %>% 
  filter(year %in% c(1906, 1956, 2006)) %>% 
  ggplot(aes(last_letter, prop)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  facet_grid(rows = vars(year)) +
  labs(
    title = "Last letter of boys' names by year",
    x = "Last letter of name",
    y = "Percentage of boys born"
  )
```

Last letter of boys' names by year.

```{r, fig.asp=1}
boys_last_letter %>% 
  ggplot(aes(year, prop)) +
  geom_line() +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  facet_wrap(facets = vars(last_letter), ncol = 5) +
  labs(
    title = "Last letter of boys' names by year",
    x = "Year",
    y = "Percentage of boys born"
  )
```

Proportion of top 10 names each year by sex.

```{r}
prop_top_10 <- 
  babynames %>% 
  group_by(year, sex) %>% 
  slice_max(order_by = prop, n = 10) %>% 
  summarize(prop = sum(prop)) %>% 
  ungroup()

prop_top_10
```

Percentage of top 10 names each year by sex.

```{r}
labels <- 
  tribble(
    ~year, ~prop, ~label,
    1948, 0.38, "Boys",
    1948, 0.18, "Girls"
  )

prop_top_10 %>% 
  ggplot(aes(year, prop)) +
  geom_line(aes(color = sex)) +
  geom_text(aes(label = label), data = labels) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  coord_cartesian(ylim = c(0, NA)) +
  theme(legend.position = "none") +
  labs(
    title = "Percentage of top 10 names each year by sex",
    x = "Year",
    y = "Percentage of babies born to each sex"
  )
```

In 2017, the last year in the data, the two lines crossed, with 7.7% for girls and 7.5% for boys.

