---
title: "Regression and Other Stories: RiskyBehavior"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Data from a randomized trial targeting couples at high risk of HIV
infection. The intervention provided counseling sessions regarding
practices that could reduce their likelihood of contracting
HIV. Couples were randomized either to a control group, a group in
which just the woman participated, or a group in which both members
of the couple participated. One of the outcomes examined after
three months was "number of unprotected sex acts." See Chapter 15
in Regression and Other Stories.

Reference: El-Bassel, N., Witte, S. S., Gilbert, L., Wu, E.,
Chang, M., Hill, J., and Steinglass, P. (2003). The efficacy of a
relationship-based HIV/STD prevention program for heterosexual
couples. *American journal of public health*, **93**, 963--969.

DOI: [10.2105/AJPH.93.6.963](https://ajph.aphapublications.org/doi/10.2105/AJPH.93.6.963)

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Data from a randomized trial targeting couples at high risk of HIV infection
file_risky <- here::here("RiskyBehavior/data/risky.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 15 Other generalized linear models

## 15.10 Exercises

```{r, message=FALSE}
risky <- read_csv(file_risky)

risky
```

