---
title: "Regression and Other Stories: Peacekeeping"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Outcomes after civil war in countries with and without United
Nations peacekeeping. See Chapter 1 in Regression and Other
Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(haven)
library(lubridate)

# Parameters
  # Outcomes after civil wars with and without U.N. peacekeeping
file_peace <- here::here("Peacekeeping/data/pk&pkept_old.dta")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 1 Overview

## 1.3 Some examples of regression

### Estimating the effects of United Nations peacekeeping, using pre-treatment variables to adjust for differences between treatment and control groups

Data

```{r}
peace <- 
  file_peace %>% 
  read_dta() %>% 
  select(id:pkop, pcw, morewar, hazard1) %>% 
  mutate(across(where(is.labelled), as_factor))

glimpse(peace)
```

Variable labels.

```{r}
peace %>% 
  keep(~ "label" %in% names(attributes(.))) %>% 
  map_chr(~ attr(., "label"))
```

Conflict ID.

```{r}
peace %>% 
  count(id, name = "nrows") %>% 
  count(nrows)
```

148 of the conflict IDs have only one row, but 22 have multiple rows. One has seven rows. Let's look at an ID with multiple rows.

```{r}
peace %>% 
  group_by(id) %>% 
  filter(n() == 7) %>% 
  select(id, cname, cfdate, faildate, t0, t1, pk_dum, pkop)
```

Haiti had a ceasefire that started on 1994-09-18. The seven rows correspond to seven different time periods. There were peacekeepers for six of these periods (`pk_dum`), though the peacekeeping operation and mission (`pkop`) varied.

We'll add a new variable `peacekeepers` to indicate whether there were peacekeepers at any time. And we'll now collapse the data to one row per conflict ID, using the first row. 

```{r}
peace <- 
  peace %>% 
  group_by(id) %>% 
  summarize(
    peacekeepers = any(pk_dum == 1),
    across(everything(), first)
  )
```

We are only interested in conflicts where we know the peacekeeper status.

```{r}
peace %>% 
  filter(is.na(peacekeepers)) %>% 
  select(id, cname, peacekeepers, cfdate, faildate, outcome, pk_dum, pkop)
```

There are 13 conflicts with no peacekeeper data. We'll remove them.

```{r}
peace <- 
  peace %>% 
  drop_na(peacekeepers)

peace %>% 
  count(peacekeepers)
```

157 conflicts remain, of which 41 had peacekeepers.

We will say that a country remained at peace if the ceasefire hadn't failed by the end of the data collection, in other words if `faildate` is `NA`. The `NA` status of `faildate` is also encoded in `morewar`.

```{r}
peace %>% 
  count(is.na(faildate), morewar)
```

The ceasefires held 72 times and failed 85 times.

Let's see the relationship between having peacekeepers and a return to conflict.

```{r}
peace %>% 
  count(peacekeepers, peace = is.na(faildate))

peace %>% 
  group_by(peacekeepers) %>% 
  summarize(peace_prop = mean(is.na(faildate)))
```

In about 51% of conflicts with peacekeepers, the ceasefires held, compared to 44% of conflicts without peacekeepers.

If conflict did return, let's see the relationship between having peacekeepers and the length of time for the return.

```{r}
peace %>% 
  filter(!is.na(faildate)) %>% 
  mutate(return_years = interval(cfdate, faildate) / dyears(1)) %>% 
  group_by(peacekeepers) %>% 
  summarize(across(return_years, list(mean = mean, median = median)))
```

In the conflicts where ceasefires did not hold, the ceasefires ended sooner when there were peacekeepers.

If conflict did not return, let's see the relationship between having peacekeepers and the duration of peace. We'll use the end of data collection, 2004-12-31, as the end date.

```{r}
peace %>% 
  filter(is.na(faildate)) %>% 
  mutate(peace_years = interval(cfdate, ymd("2004-12-31")) / dyears(1)) %>% 
  group_by(peacekeepers) %>% 
  summarize(across(peace_years, list(mean = mean, median = median)))
```

In the conflicts where the ceasefires held, the duration of peace was shorter when there were peacekeepers.

Years until return of war.

```{r}
ceasefire_failed <- 
  peace %>% 
  filter(!is.na(faildate)) %>% 
  mutate(return_years = interval(cfdate, faildate) / dyears(1))

return_years_means <- 
  ceasefire_failed %>% 
  group_by(peacekeepers) %>% 
  summarize(return_years_mean = mean(return_years))

peacekeepers_labels <- 
  c(
    "FALSE" = "Without peacekeeping",
    "TRUE" = "With peacekeeping"
  )

ceasefire_failed %>% 
  ggplot(aes(return_years)) +
  geom_histogram(binwidth = 0.5, boundary = 0) +
  geom_vline(aes(xintercept = return_years_mean), data = return_years_means) +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  facet_grid(
    rows = vars(peacekeepers),
    labeller = labeller(peacekeepers = peacekeepers_labels)
  ) +
  labs(
    title = "Years until return of war",
    x = "Years",
    y = "Count"
  )
```

Years until return of war.

```{r}

ceasefire_failed %>% 
  drop_na(hazard1) %>% 
  ggplot(aes(hazard1, return_years, color = peacekeepers)) +
  geom_point() +
  scale_x_log10(
    breaks = c(1.4e-5, 4.5e-2),
    minor_breaks = NULL,
    labels = c("Better conditions", "Worse conditions")
  ) +
  scale_y_continuous(breaks = scales::breaks_width(5)) +
  scale_color_discrete(
    breaks = c(FALSE, TRUE),
    labels = c("Without peacekeeping", "Without peacekeeping")
  ) +
  theme(
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    panel.grid.major.x = element_blank()
  ) +
  labs(
    title = "Years until return of war",
    x = "Pre-treatment measure of conditions in the country",
    y = "Years",
    color = NULL
  )
```

The plot uses `hazard1` as a measure of the conditions in the country at the time of the ceasefire. Below a threshold of this measure, it was possible for ceasefires to last 10 years or longer. Over this threshold, all of the countries returned to war in 5 years or less.

