---
title: "Regression and Other Stories: Poststratification"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Poststratification after estimation. See Chapter 17 in Regression
and Other Stories.

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 17 Poststratification and missing-data imputation

## 17.1 Poststratification: using regression to generalize to a new population

### Adjusting for a single factor

The CBS News poll conducted from 12 -- 16 October 2016 reported that,among likely voters who preferred one of the two major-party candidates, 45% intended to vote for Donald Trump and 55% for Hillary Clinton. Of these respondents, 33% reported Republican party
identification, 40% affiliated themselves with Democrats, and 27% did not declare a major-party affiliation.

Sources:

* [CBS poll: Clinton's lead over Trump widens with three weeks to go](http://www.cbsnews.com/news/cbs-poll-clintons-lead-over-trump-widens-with-three-weeks-to-go/). 2016-10-17.
* [CBS New Poll 10-17 toplines](https://www.scribd.com/document/327938789/CBS-News-Poll-10-17-toplines). 2016-10-17.

Data

```{r}
exit_poll_2016 <- 
  tribble(
    ~party,       ~exit_pct,
    "Republican",        33,
    "Democrat",          36,
    "Independent",       31
  )

poll_n <- 
  tribble(
    ~party,         ~n,
    "Republican",  254,
    "Democrat",    282,
    "Independent", 242
  )

poll_stats <- 
  tribble(
    ~party,        ~candidate, ~vote_pct,
    "Republican",  "Trump",           77,
    "Republican",  "Clinton",          8,
    "Republican",  NA,                15,
    "Democrat",    "Trump",            5,
    "Democrat",    "Clinton",         89,
    "Democrat",    NA,                 6,
    "Independent", "Trump",           36,
    "Independent", "Clinton",         38,
    "Independent", NA,                26
  )
```

Check that survey results add up to 100%.

```{r}
poll_stats %>% 
  group_by(party) %>% 
  summarize(vote_pct = sum(vote_pct))
```

Create dataset with above survey response statistics.

```{r}
poll <- 
  poll_stats %>% 
  left_join(poll_n, by = "party") %>% 
  mutate(vote = round(n * vote_pct / 100)) %>% 
  group_by(party) %>% 
  mutate(
    vote = 
      if_else(!is.na(candidate), vote, first(n) - sum(vote[!is.na(candidate)]))
  ) %>% 
  ungroup() %>% 
  mutate(vote = map2(candidate, vote, rep)) %>% 
  select(party, vote) %>% 
  unnest(vote)

poll
```

Check that `poll` has above survey response statistics.

```{r}
poll %>% 
  count(party) %>% 
  all.equal(poll_n %>% arrange(party))

poll %>% 
  count(party, vote) %>% 
  group_by(party) %>% 
  mutate(vote_pct = round(100 * n / sum(n))) %>% 
  ungroup() %>% 
  select(party, candidate = vote, vote_pct) %>% 
  arrange(party, candidate) %>% 
  all.equal(poll_stats %>% arrange(party, candidate))
```

#### Poststratification using a weighted average

```{r}
poll_stats %>% 
  drop_na(candidate) %>% 
  group_by(party) %>% 
  mutate(vote_prop = vote_pct / sum(vote_pct)) %>% 
  ungroup() %>% 
  filter(candidate == "Trump") %>% 
  left_join(exit_poll_2016, by = "party") %>% 
  summarize(trump_prop = weighted.mean(vote_prop, exit_pct))
```

#### What if we had weighted using the distribution of party identification in the sample?

```{r}
mean(poll$vote == "Trump", na.rm = TRUE)
```

#### Poststratification using regression prediction

Fit linear model to adjust for just one factor, party identification.

```{r}
set.seed(660)

poll <- 
  poll %>% 
  mutate(vote_trump = vote == "Trump")

fit_1 <- stan_glm(vote_trump ~ party, data = poll, refresh = 0)

print(fit_1, digits = 2)
```

We set up a poststratification matrix.

```{r}
poststrat_1 <- 
  exit_poll_2016 %>% 
  transmute(
    party,
    prop = exit_pct / sum(exit_pct)
  )

poststrat_1
```

Use the fitted model to make predictions for each row of the poststratification matrix.

```{r}
epred_1 <- posterior_epred(fit_1, newdata = poststrat_1)

dim(epred_1)
```

The result is a `r nrow(epred_1)` x `r ncol(epred_1)` matrix representing the posterior uncertainty in the proportion of Trump supporters in each of the three strata.

We sum these to get the population average.

```{r}
poststrat_est <- tibble(trump_prop = epred_1 %*% poststrat_1$prop)

poststrat_est %>% 
  summarize(across(trump_prop, list(mean = mean, mad = mad)))
```

Suppose, for example, we would like to say that political polls can easily be off by 2 percentage points, merely due to nonresponse and changes of opinion. We could then add this uncertainty to our prediction:

```{r}
set.seed(274)

poll_error <- 0.02

poststrat_est %>% 
  mutate(trump_prop = trump_prop + rnorm(n(), mean = 0, sd = poll_error)) %>% 
  summarize(across(trump_prop, list(mean = mean, mad = mad)))
```

This new estimate has the same mean but a larger standard error.

