---
title: "Regression and Other Stories: Robit"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Comparison of robit and logit models for binary data. See Chapter
15 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(cmdstanr)

# Parameters
  # Stan logistic model
file_logit <- here::here("Robit/logit_tv.stan")
  # Stan robit model
file_robit <- here::here("Robit/robit_tv.stan")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 15 Other generalized linear models

## 15.6 Robust regression using the t model

### Robit instead of logit or probit

Simulate clean data.

```{r}
set.seed(1234)

N <- 50
df <- 4

x_min <- -9
x_max <- 9
a <- 0
b <- 0.8

data_clean <- 
  tibble(
    x = runif(N, min = x_min, max = x_max),
    y = rbinom(N, 1, prob = plogis(a + b * x))
  )

data_1 <- c(list(N = N, df = df), as.list(data_clean))
```

Stan logistic model.

```{r, comment=""}
file_logit %>% 
  read_lines() %>% 
  writeLines()
```

Stan robit model.

```{r, comment=""}
file_robit %>% 
  read_lines() %>% 
  writeLines()
```

Compile models.

```{r, message=FALSE}
model_logit <- cmdstan_model(file_logit)
model_robit <- cmdstan_model(file_robit)
```

Sample from logistic model with clean data.

```{r}
set.seed(693)

fit_1_logit <- 
  model_logit$sample(data = data_1, refresh = 0, adapt_delta = 0.95)

fit_1_logit
```

Logistic regression coefficients.

```{r}
a_1_logit <- median(fit_1_logit$draws("a"))
b_1_logit <- median(fit_1_logit$draws("b"))
```

Sample from robit model with clean data.

```{r}
set.seed(693)

fit_1_robit <- 
  model_robit$sample(data = data_1, refresh = 0, adapt_delta = 0.95)

fit_1_robit
```

Robit regression coefficients.

```{r}
a_1_robit <- median(fit_1_robit$draws("a"))
b_1_robit <- median(fit_1_robit$draws("b"))
```

Logistic and robit regressions with clean data.

```{r, fig.asp=0.75}
lines <- function(a_logit, b_logit, a_robit, b_robit) {
   tibble(
    x = seq_range(c(x_min, x_max)),
    y_logit = plogis(a_logit + b_logit * x),
    y_robit = pt((a_robit + b_robit * x) / sqrt((df - 2) / df), df = df)
  ) %>% 
    pivot_longer(
      cols = !x,
      names_to = "model",
      names_prefix = "y_",
      values_to = "y"
    )
}

lines_clean <- lines(a_1_logit, b_1_logit, a_1_robit, b_1_robit)

plot <- function(data, lines) {
  ggplot(mapping = aes(x, y)) +
    geom_point(data = data, alpha = 0.75) +
    geom_line(aes(color = model), data = lines) +
    scale_y_continuous(breaks = 0:1, minor_breaks = NULL) +
    scale_color_discrete(
      breaks = c("logit", "robit"),
      labels = c("Logistic regression", "Robit regression")
    ) +
    theme(legend.position = "bottom") +
    labs(color = NULL)
}

plot(data_clean, lines_clean) +
  labs(title = "Logistic and robit regressions with clean data")
```

Add an outlier by flipping the class of one observation.

```{r}
data_contaminated <- 
  data_clean %>% 
  arrange(x) %>% 
  mutate(y = if_else(row_number() == 4, 1L, y))

data_2 <- c(list(N = N, df = df), as.list(data_contaminated))
```

Sample from logistic model with contaminated data.

```{r}
set.seed(693)

fit_2_logit <- 
  model_logit$sample(data = data_2, refresh = 0, adapt_delta = 0.95)

fit_2_logit
```

Logistic regression coefficients.

```{r}
a_2_logit <- median(fit_2_logit$draws("a"))
b_2_logit <- median(fit_2_logit$draws("b"))
```

Sample from robit model with contaminated data.

```{r}
set.seed(693)

fit_2_robit <- 
  model_robit$sample(data = data_2, refresh = 0, adapt_delta = 0.95)

fit_2_robit
```

Robit regression coefficients.

```{r}
a_2_robit <- median(fit_2_robit$draws("a"))
b_2_robit <- median(fit_2_robit$draws("b"))
```

Logistic and robit regressions with contaminated data.

```{r, fig.asp=0.75}
lines_contaminated <- lines(a_2_logit, b_2_logit, a_2_robit, b_2_robit)

plot(data_contaminated, lines_contaminated) +
  labs(title = "Logistic and robit regressions with contaminated data")
```

