---
title: "Regression and Other Stories: Sample size simulation"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Sample size simulation. See Chapter 16 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Seed
SEED <- 660
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 16 Design and sample size decisions

## 16.4 Interactions are harder to estimate than main effects

### Understanding the problem by simulating regressions in R

#### Simulated data 1: predictor range: -0.5, 0.5

```{r}
set.seed(SEED)

n <- 1000
sigma <- 10

y <- rnorm(n, mean = 0, sd = sigma)
sample_1 <- rep(1:2, n / 2) %>%  sample()
sample_2 <- rep(1:2, n / 2) %>%  sample()

sim <- function(v_1, v_2) {
  tibble(
    y,
    x_1 = c(v_1, v_2)[sample_1],
    x_2 = c(v_1, v_2)[sample_2]
  )
}

data_1 <- sim(-0.5, 0.5)
```

Fit model with one predictor.

```{r}
fit_1_1 <- stan_glm(y ~ x_1, data = data_1, refresh = 0, seed = SEED)

fit_1_1
```

Fit model with two predictors with an interaction.

```{r}
fit_1_2 <- 
  stan_glm(y ~ x_1 + x_2 + x_1:x_2, data = data_1, refresh = 0, seed = SEED)

fit_1_2
```

Ignore the estimates; they're pure noise. Just look at the standard errors. They follow the formulas:

Main effect:

```{r}
2 * sigma / sqrt(n)
```

Interaction:

```{r}
4 * sigma / sqrt(n)
```

#### Simulated data 2: predictor range: 0, 1

```{r}
data_2 <- sim(0, 1)
```

Fit model with one predictor.

```{r}
fit_2_1 <- stan_glm(y ~ x_1, data = data_2, refresh = 0, seed = SEED)

fit_2_1
```

Fit model with two predictors with an interaction.

```{r}
fit_2_2 <- 
  stan_glm(y ~ x_1 + x_2 + x_1:x_2, data = data_2, refresh = 0, seed = SEED)

fit_2_2
```

In this case, the standard error for the interaction is unchanged, and the standard error for the main effects have increased by a factor of sqrt(2).

#### Simulated data 3: predictor range: -1, 1

```{r}
data_3 <- sim(-1, 1)
```

Fit model with one predictor.

```{r}
fit_3_1 <- stan_glm(y ~ x_1, data = data_3, refresh = 0, seed = SEED)

fit_3_1
```

Fit model with two predictors with an interaction.

```{r}
fit_3_2 <- 
  stan_glm(y ~ x_1 + x_2 + x_1:x_2, data = data_3, refresh = 0, seed = SEED)

fit_3_2
```

In this case, multiplying the predictors by 2 has the effect of dividing the standard errors for the main effects by 2 and the standard errors for the interaction by 4.

