---
title: "Regression and Other Stories: Congress"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Predictive uncertainty for congressional elections. See Chapter 10
in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Congressional election data
file_congress <- here::here("Congress/data/congress.csv")
  # Party colors
party_colors <- 
  c(
    "Democrat"   = "#1a80c4",
    "Republican" = "#cc3d3d",
    "Open"       = "#7ead53"
  )
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 10 Linear regression with multiple predicators

## 10.6 Example: uncertainty in predicting congressional elections

### Background

Data

```{r, message=FALSE}
congress <- 
  file_congress %>% 
  read_csv() %>% 
  mutate(
    across(
      starts_with("inc"),
      ~ case_when(
        . == -1 ~ "Republican",
        . ==  0 ~ "Open",
        . ==  1 ~ "Democrat",
        TRUE ~ NA_character_
      )
    )
  )

congress
```

The `inc*` variables represent whether an incumbent is running for reelection and, if so, their party.

The `*_adj` variables represent adjustments to account for uncontested elections. If `vx` is 0, then `vx_adj` is 0.25. If `vx` is greater than 0.9, then `vx_adj` is 0.75.

Congressional elections in 1988: Raw data.

```{r}
congress %>% 
  ggplot(aes(v88)) +
  geom_histogram(binwidth = 0.05, boundary = 0) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(
    title = "Congressional elections in 1988",
    subtitle = "Raw data",
    x = "Democratic share of two-party vote",
    y = "Count"
  )
```

Congressional elections in 1986 and 1988: Raw data.

```{r, fig.asp=1}
set.seed(616)

congress %>% 
  {
    bind_rows(
      filter(., inc88 != "Open"),
      filter(., inc88 == "Open")
    )
  } %>% 
  ggplot(aes(v86, v88, color = inc88)) +
  geom_hline(yintercept = 0.5, color = "grey60") +
  geom_vline(xintercept = 0.5, color = "grey60") +
  geom_abline(slope = 1, intercept = 0) +
  geom_count() +
  coord_fixed() +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = party_colors) +
  guides(size = "none") +
  labs(
    title = "Congressional elections in 1986 and 1988",
    subtitle = "Raw data",
    x = "Democratic vote share in 1986",
    y = "Democratic vote share in 1988",
    color = "Incumbent\nin 1988"
  )
```

### Data issues

Congressional elections in 1986 and 1988: Adjusted data.

```{r, fig.asp=1}
set.seed(616)

congress %>% 
  {
    bind_rows(
      filter(., inc88 != "Open"),
      filter(., inc88 == "Open")
    )
  } %>% 
  ggplot(aes(v86_adj, v88_adj, color = inc88)) +
  geom_hline(yintercept = 0.5, color = "grey60") +
  geom_vline(xintercept = 0.5, color = "grey60") +
  geom_abline(slope = 1, intercept = 0) +
  geom_count() +
  coord_fixed(xlim = 0:1, ylim = 0:1) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = party_colors) +
  guides(size = "none") +
  labs(
    title = "Congressional elections in 1986 and 1988",
    subtitle = "Adjusted data",
    x = "Democratic vote share in 1986",
    y = "Democratic vote share in 1988",
    color = "Incumbent\nin 1988"
  )
```

### Fitting the model

Linear regression to predict `vote` (the Democratic share of the two-party vote in each district), given `past_vote` (the Democrats' share in the previous election) and `incumbent` (incumbency of seat).

```{r}
set.seed(905)

data_88 <-  
  congress %>% 
  transmute(
    vote = v88_adj,
    vote_prev = v86_adj,
    incumbent = inc88
  )

fit_88 <- stan_glm(vote ~ vote_prev + incumbent, data = data_88, refresh = 0)

print(fit_88, digits = 2)
```

### Simulation for inferences and predictions of new data points

Running `stan_glm()` produces a set of simulation draws expressing uncertainty in the parameters in the fitted model. We can access these simulations by extracting them from the fitted model object:

```{r}
sims_88 <- as_tibble(fit_88)

nrow(sims_88)
```

We can use these simulations, along with data from 1988 and incumbency information in 1990, to predict the district-by-district election outcome in 1990. We start by creating a new tibble of predictors:

```{r}
data_90 <-  
  congress %>% 
  transmute(
    vote_prev = v88_adj,
    incumbent = inc90
  )

nrow(data_90)
```

We then simulate predictive simulations of new outcomes:

```{r}
set.seed(620)

pred_90 <- 
  posterior_predict(fit_88, newdata = data_90) %>% 
  as_tibble()

dim(pred_90)
```

The resulting tibble has `r nrow(pred_90)` rows, one for each simulation, and `r ncol(pred_90)` columns, one for each predicted congressional district.

### Predictive simulation for a nonlinear function of new data

For the congressional elections example, to perform inference on the predicted number of elections won by the Democrats in 1990, we sum over the rows of the tibble:

```{r}
pred_90_dems <- 
  pred_90 %>% 
  mutate(across(everything(), ~ . > 0.5)) %>% 
  rowwise() %>% 
  mutate(pred_dems = sum(c_across(everything()))) %>% 
  pull(pred_dems)
```

Predicted number of Democratic wins in 1990.

```{r}
tibble(pred_90_dems = pred_90_dems) %>% 
  ggplot(aes(pred_90_dems)) +
  geom_bar() +
  labs(
    title = "Predicted number of Democratic wins in 1990",
    x = "Predicted number of Democratic wins",
    y = "Count"
  )
```

```{r, comment=""}
cat(
  str_glue(
    "The mean predicted number of Democratic wins in 1990 is ",
    "{format(mean(pred_90_dems), digits = 1, nsmall = 1)} with a standard ",
    "deviation of {format(sd(pred_90_dems), digits = 1, nsmall = 1)}."
  )
)
```

The actual number of Democratic wins in 1990 was 267.

```{r}
actual_90_dems <- 267

z <- (actual_90_dems - mean(pred_90_dems)) / sd(pred_90_dems)
z
```

This was `r format(z, digits = 2, nsmall = 2)` standard deviations from the predicted number of wins.

