---
title: "Regression and Other Stories: Congress"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r format(Sys.Date())`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Predictive uncertainty for congressional elections. See Chapter 2
in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Directory with congressional election data
dir_elections <- here::here("Congress/data")
  # Data variables
elections_rename <- 
  c(
    state_code = "X1",
    district_code = "X2",
    incumbent = "X3",
    d_vote = "X4",
    r_vote = "X5"
  )
  # Regions
regions <- c("Northeast", "Midwest", "South", "West")
  # Common code
file_common <- here::here("_common.R")
  
#===============================================================================

# Run common code
source(file_common)
```

# 2 Data and measurement

## 2.3 All graphs are comparisons

### Grids of plots

Data

```{r, message=FALSE}
elections <- 
  fs::dir_ls(path = dir_elections, regexp = "\\d{4}.asc$") %>% 
  map_dfr(~ read_table2(., col_names = FALSE), .id = "year") %>% 
  rename(!!! elections_rename) %>% 
  mutate(
    year = str_match(year, "(\\d{4}).asc$")[, 2] %>% as.integer(),
    across(!year, na_if, -9),
    incumbent =
      case_when(
        incumbent %in% c(-1, 1) ~ TRUE,
        incumbent == 0 ~ FALSE,
        TRUE ~ NA
      )
  )
```

```{r}
elections %>% 
  count(is.na(d_vote), is.na(r_vote))
```

We'll omit races with `NA`s for votes.

```{r}
elections <- 
  elections %>% 
  drop_na(d_vote, r_vote)
```

```{r}
range(elections$state_code)

elections %>% 
  count(state_code) %>% 
  slice_max(order_by = state_code, n = 5)
```

The formula in the original code to assign regions is `floor(state_code / 20) + 1`. Since there are only four regions, we will omit rows with `state_code` 81 and 82, for which the formula would yield 5.

```{r}
elections <- 
  elections %>% 
  filter(!state_code %in% 81:82) %>% 
  mutate(region = regions[floor(state_code / 20) + 1])
```

Finally, we'll calculate the Democratic proportion of the vote for the current and previous election.

```{r}
elections <- 
  elections %>% 
  mutate(d_prop = d_vote / (d_vote + r_vote))

elections <- 
  sort(unique(elections$year))[-1] %>% 
  map_dfr(
    ~ elections %>% 
      filter(year == .x) %>% 
      inner_join(
        elections %>% 
          filter(year == .x - 2) %>% 
          select(state_code, district_code, d_prop_prev = d_prop),
        by = c("state_code", "district_code")
      )
  )

elections %>% 
  slice(1:10) %>% 
  knitr::kable()
```

Swings in U.S. congressional elections.

```{r, fig.asp=0.8, fig.width=10}
v <- 
  elections %>% 
  filter(
    year %in% c(1950, 1970, 1990),
    !is.na(incumbent),
    abs(d_prop - 0.5) < 0.3 & abs(d_prop_prev - 0.5) < 0.3
  ) %>% 
  mutate(period = str_glue("{year - 2} to {year}")) %>% 
  {
    bind_rows(
      filter(., incumbent),
      filter(., !incumbent)
    )
  }

v %>% 
  ggplot() +
  geom_vline(xintercept = 0.5, color = "grey60") +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_point(aes(d_prop_prev, d_prop - d_prop_prev, color = incumbent)) +
  facet_grid(rows = vars(period), cols = vars(region)) +
  coord_fixed() +
  scale_x_continuous(
    breaks = scales::breaks_width(0.1),
    minor_breaks = NULL,
    labels = scales::label_percent(accuracy = 1),
    expand = expansion(add = 0.05)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1),
    minor_breaks = NULL,
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_color_discrete(
    breaks = c(FALSE, TRUE),
    labels = c("Open seat", "Incumbent running")
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Swings in U.S. congressional elections",
    subtitle = "Where Democratic percentage was between 20 - 80%",
    x = "Democratic percentage in first election",
    y = "Swing in Democratic vote in next election",
    color = NULL
  )
```

