---
title: "Regression and Other Stories: Roaches"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Analyze the effect of integrated pest management on reducing
cockroach levels in urban apartments. See Chapter 15 in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(bayesplot)
library(brms)
library(rstanarm)

# Parameters
  # Seed
SEED <- 3579
  # Common code
file_common <- here::here("_common.R")

# Functions
  # Plot kernel density of data and sample replicates
plot_density_overlay <- function(y, y_rep) {
  ggplot(mapping = aes(y)) +
    stat_density(
      aes(group = rep, color = "y_rep"),
      data = 
        seq_len(nrow(y_rep)) %>% map_dfr(~ tibble(rep = ., y = y_rep[., ])),
      geom = "line",
      position = "identity",
      alpha = 0.5,
      size = 0.25
    ) +
    stat_density(aes(color = "y"), data = tibble(y), geom = "line", size = 1) +
    scale_y_continuous(breaks = 0) +
    scale_color_discrete(
      breaks = c("y", "y_rep"),
      labels = c("y", expression(y[rep]))
    ) +
    theme(legend.text.align = 0) +
    labs(
      x = NULL,
      y = NULL,
      color = NULL
    )
}

#===============================================================================

# Run common code
source(file_common)
```

# 15 Other generalized linear models

## 15.2 Poisson and negative binomial regression

### Example: zeros in count data

Data

```{r}
# data(roaches)
roaches <-
  roaches %>%
  as_tibble() %>% 
  mutate(roach100 = roach1 / 100) %>%
  relocate(roach100, .after = roach1)

roaches
```

Fit negative binomial model.

```{r}
fit_nbinom <-
  stan_glm(
    y ~ roach100 + treatment + senior,
    family = neg_binomial_2(link = "log"),
    data = roaches,
    offset = log(exposure2),
    refresh = 0,
    seed = SEED
  )

print(fit_nbinom, digits = 2)
```

The estimated overdispersion is 1 / `reciprocal_dispersion` = 1 / 0.27 = 3.7. The model is overdispersed compared to Poisson.

LOO log score

```{r}
loo_nbinom <- loo(fit_nbinom)
```

Because of the warning message, we will call `loo()` again with the recommended argument.

```{r, message=FALSE}
loo_nbinom <- loo(fit_nbinom, k_threshold = 0.7)

loo_nbinom
```

#### Checking model fit by comparing the data, `y`, to the replicated datasets, $y^{rep}$

Simulate data for negative binomial model.

```{r}
set.seed(SEED)

y_rep_nbinom <- posterior_predict(fit_nbinom)

n_sims <- nrow(y_rep_nbinom)
n_rep <- 100
sims_sample <- sample(n_sims, n_rep)
```

Kernel density of data and `r n_rep` sample replicates from negative binomial model.

```{r}
plot_density_overlay(
  y = log10(roaches$y + 1),
  y_rep = log10(y_rep_nbinom[sims_sample, ] + 1)
) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = 
      str_glue(
        "Kernel density of data and {n_rep} sample replicates from negative binomial model"
      ),
    x = "log10(y + 1)"
  )
```

Kernel density of data and `r n_rep` sample replicates from negative binomial model using bayesplot.

```{r}
ppc_dens_overlay(
  y = log10(roaches$y + 1),
  yrep = log10(y_rep_nbinom[sims_sample, ] + 1)
) +
  theme(
    axis.line.y = element_blank(),
    text = element_text(family = "sans")
  ) +
  labs(
    title = "Negative binomial model",
    x = "log10(y + 1)"
  )
```

##### Predictive checking with test statistic

###### Proportion of zero counts

Test statistic for proportion of zero counts.

```{r}
prop_0 <- function(x) {
  mean(x == 0)
}
```

```{r}
y_prop_0 <- prop_0(roaches$y)

y_prop_0
```

```{r}
y_rep_nbinom_prop_0 <- apply(y_rep_nbinom, 1, prop_0)

summary(y_rep_nbinom_prop_0)
```

The proportion of zero counts in the replicated datasets varies from 0.21 to 0.49, with a median and mean of 0.34, close to the observed proportion in the data of 0.36.

Distribution of proportion of zero counts in replicates.

```{r, fig.asp=0.75}
tibble(y_prop_0 = y_rep_nbinom_prop_0) %>% 
  ggplot(aes(y_prop_0)) +
  geom_histogram(binwidth = 0.015) +
  geom_vline(xintercept = y_prop_0, color = "red") +
  labs(
    title = "Distribution of proportion of zero counts in replicates",
    subtitle = "Vertical line is proportion of zero counts in data",
    x = "Proportion of zero counts in replicate",
    y = "Count"
  )
```

Distribution of proportion of zero counts in replicates using bayesplot.

```{r, message=FALSE}
ppc_stat(y = roaches$y, yrep = y_rep_nbinom, stat = function(x) mean(x == 0))
```

###### Maximum counts

```{r}
y_max <- max(roaches$y)

y_max
```

```{r}
y_rep_nbinom_max <- apply(y_rep_nbinom, 1, max)

summary(y_rep_nbinom_max)
```

The maximum roach count in the data is `r y_max`, but the maximum count in the negative binomial model is close to nine million, which would make sense only in a horror movie. We will introduce another improvement below with the zero-inflated negative binomial model.

#### What if we had used Poisson regression?

Fit Poisson model.

```{r}
fit_pois <-
  stan_glm(
    y ~ roach100 + treatment + senior,
    family = poisson(link = "log"),
    data = roaches,
    offset = log(exposure2),
    refresh = 0,
    seed = SEED
  )

print(fit_pois, digits = 2)
```

LOO log score

```{r}
loo_pois <- loo(fit_pois)
```

Because of the warning message, we will follow the recommendation and use `kfold()` instead of `loo()`.

K-fold log score

```{r, message=FALSE}
kfold_pois <- kfold(fit_pois, k = 10)

kfold_pois
```

Compare log scores.

```{r, message=FALSE}
kfold_nbinom <- kfold(fit_nbinom, k = 10)

loo_compare(kfold_nbinom, kfold_pois)
```

The negative binomial model has a substantially better log score than the Poisson model.

Simulate data for Poisson model.

```{r}
set.seed(SEED)

y_rep_pois <- posterior_predict(fit_pois)
```

Kernel density of data and `r n_rep` sample replicates from Poisson model.

```{r}
plot_density_overlay(
  y = log10(roaches$y + 1),
  y_rep = log10(y_rep_pois[sims_sample, ] + 1)
) +
  labs(
    title = 
      str_glue(
        "Kernel density of data and {n_rep} sample replicates from Poisson model"
      ),
    x = "log10(y + 1)"
  )
```

Kernel density of data and `r n_rep` sample replicates from Poisson model using bayesplot.

```{r}
ppc_dens_overlay(
  y = log10(roaches$y + 1),
  yrep = log10(y_rep_pois[sims_sample, ] + 1)
) +
  theme(
    axis.line.y = element_blank(),
    text = element_text(family = "sans")
  ) +
  labs(
    title = "Poisson model",
    x = "log10(y + 1)"
  )
```

#### Checking the fit of the non-overdispersed Poisson regression

Test statistic for proportion of zero counts.

```{r}
y_rep_pois_prop_0 <- apply(y_rep_pois, 1, prop_0)

summary(y_rep_pois_prop_0)
```

A majority of the replicated datasets have no zero counts, the maximum proportion in all `r n_sims` replicates is 0.0076, and the mean proportion is 0.00070, far below the observed proportion in the data of 0.36. This suggests a potential problem with the model: in reality, many apartments have zero roaches, but this would not be happening if the model were true.

## 15.8 Going beyond generalized linear models

### Cockroaches and the zero-inflated negative binomial model

A zero-inflated negative binomial model is a mixture of two models, a logistic regression to model the proportion of extra zero counts and a negative binomial model to model the non-zero counts. This two-stage model can be modeled directly in Stan or with `brms::brm()` using the `zero_inflated_negbinomial()` family. We'll use the latter here.

Create log-transformed variables.

```{r}
roaches <- 
  roaches %>% 
  mutate(
    roach1_logp1 = log(roach1 + 1),
    exposure2_log = log(exposure2)
  )
```

Fit zero-inflated negative binomial model.

```{r, message=FALSE, warning=FALSE, error=FALSE, results=FALSE}
fit_zinbinom <- 
  brm(
    formula = 
      bf(
        y  ~ roach1_logp1 + treatment + senior + offset(exposure2_log),
        zi ~ roach1_logp1 + treatment + senior + offset(exposure2_log)
      ),
    data = roaches,
    family = zero_inflated_negbinomial(),
    prior = set_prior("normal(0, 1)"),
    save_all_pars = TRUE,
    seed = SEED,
    refresh = 0
  )
```

```{r}
fit_zinbinom
```

LOO log score

```{r}
loo_zinbinom <- loo(fit_zinbinom)
```

Because of the warning message, we will call `loo()` again with the recommended argument.

```{r}
loo_zinbinom <- loo(fit_zinbinom, moment_match = TRUE)

loo_zinbinom
```

Compare log scores.

```{r}
loo_compare(loo_nbinom, loo_zinbinom)
```

The LOO log score comparison supports better predictive performance for the zero-inflated negative binomial model.

Simulate data for zero-inflated negative binomial model.

```{r}
set.seed(SEED)

y_rep_zinbinom <- posterior_predict(fit_zinbinom)
```

Kernel density of data and `r n_rep` sample replicates from zero-inflated negative binomial model.

```{r}
plot_density_overlay(
  y = log10(roaches$y + 1),
  y_rep = log10(y_rep_zinbinom[sims_sample, ] + 1)
) +
  labs(
    title = 
      str_glue(
        "Kernel density of data and {n_rep} sample replicates\nfrom zero-inflated negative binomial model"
      ),
    x = "log10(y + 1)"
  )
```

The zero-inflated negative binomial model predictions are visually better than with the negative binomial model.

Kernel density of data and `r n_rep` sample replicates from zero-inflated negative binomial model using bayesplot.

```{r}
ppc_dens_overlay(
  y = log10(roaches$y + 1),
  yrep = log10(y_rep_zinbinom[sims_sample, ] + 1)
) +
  theme(
    axis.line.y = element_blank(),
    text = element_text(family = "sans")
  ) +
  labs(
    title = "Zero-inflated negative binomial model",
    x = "log10(y + 1)"
  )
```

##### Predictive checking with test statistic

###### Proportion of zero counts

Test statistic for proportion of zero counts.

```{r}
y_rep_zinbinom_prop_0 <- apply(y_rep_zinbinom, 1, prop_0)

summary(y_rep_zinbinom_prop_0)
```

The proportion of zero counts in the replicated datasets have a median and mean of closer to the observed proportion in the data of 0.36.

Distribution of proportion of zero counts in replicates.

```{r, fig.asp=0.75}
tibble(y_prop_0 = y_rep_zinbinom_prop_0) %>% 
  ggplot(aes(y_prop_0)) +
  geom_histogram(binwidth = 0.015) +
  geom_vline(xintercept = y_prop_0, color = "red") +
  labs(
    title = "Distribution of proportion of zero counts in replicates",
    subtitle = "Vertical line is proportion of zero counts in data",
    x = "Proportion of zero counts in replicate",
    y = "Count"
  )
```

Distribution of proportion of zero counts in replicates using bayesplot.

```{r, message=FALSE}
ppc_stat(y = roaches$y, yrep = y_rep_zinbinom, stat = function(x) mean(x == 0))
```

###### Maximum counts

```{r}
y_rep_zinbinom_max <- apply(y_rep_zinbinom, 1, max)

summary(y_rep_zinbinom_max)
```

The maximum count in the data is `r y_max`. The maximum count in the replicates for the zero-inflated negative binomial model is much closer to this value than the value of nearly nine million for the negative binomial model.

