---
title: "Regression and Other Stories: Coverage"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Coverage - Illustration of coverage of intervals. See Chapter 4 in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 4 Statistical inference

## 4.2 Estimates, standard errors, and confidence intervals

### Standard errors, inferential uncertainty, and confidence intervals

Simulated data.

```{r}
set.seed(224)

n_sim <- 100
mean <- 6
sd <- 4

sims <- 
  tibble(
    sim = seq_len(n_sim),
    x = rnorm(n_sim, mean = mean, sd = sd),
    lower_95 = x + qnorm(0.025, mean = 0, sd = sd),
    lower_50 = x + qnorm(0.25, mean = 0, sd = sd),
    upper_50 = x + qnorm(0.75, mean = 0, sd = sd),
    upper_95 = x + qnorm(0.975, mean = 0, sd = sd)
  )
```

Coverage.

```{r}
v <- 
  sims %>% 
  summarize(
    coverage_50 = mean(mean >= lower_50 & mean <= upper_50),
    coverage_95 = mean(mean >= lower_95 & mean <= upper_95)
  )

v
```

The 50% uncertainty interval contained the simulation population mean `r format(100 * v$coverage_50, digits = 0)`% of the time, and the 95% uncertainty interval contained the simulation population mean `r format(100 * v$coverage_95, digits = 0)`% of the time.

Coverage of `r n_sim` random simulations.

```{r, fig.asp=0.75}
sims %>% 
  ggplot(aes(sim)) +
  geom_hline(yintercept = mean, color = "red") +
  geom_linerange(aes(ymin = lower_95, ymax = upper_95)) +
  geom_linerange(aes(ymin = lower_50, ymax = upper_50), size = 1) +
  geom_point(aes(y = x)) +
  labs(
    title = str_glue("Coverage of {n_sim} random simulations"),
    subtitle = "True value in red",
    x = "Simulation",
    y = "Estimates with 50% and 95% uncertainty intervals"
  )
```

